---
title: "This pipeline to get nhoods data from the simulations for DA-DE"
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---
# Load dependencies

```{r load, message = FALSE}


library(SingleCellExperiment)
library(BiocParallel)
library(geneBasisR)
library(splatter)
library(scran)
library(miloR)
library(miloDE)
library(stringr)
library(ggplot2)
library(ggpubr)
library(viridis)
library(MetBrewer)
library(dplyr)
library(tibble)

ncores = 24
mcparam = MulticoreParam(workers = ncores)
register(mcparam)
set.seed(32)

root.dir = "/nfs/research/marioni/alsu/hubmap_metaRef/"
source(paste0(root.dir , "miloDE_analysis/core_functions.R"))
figures.dir = paste0(root.dir , "figures/da_de/simulations_splatter/random/")

# generated simulation data
sces = readRDS(file = paste0(root.dir , "data/processed/da_de/simulations_random/sims.Rds"))
simulated_genes = readRDS(file = paste0(root.dir , "data/processed/da_de/simulations_random/simulated_genes.Rds"))
stat_da = readRDS(file = paste0(root.dir , "data/processed/da_de/simulations_random/stat_da.Rds"))
stat_de = readRDS(file = paste0(root.dir , "data/processed/da_de/simulations_random/stat_de.Rds"))

stat_combined = readRDS(file = paste0(root.dir , "data/processed/da_de/simulations_random/stat_combined_all_genes.Rds"))


```


# Heatmaps

## DA

### Split by all

```{r heatmap-da-split-all, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "frac_hoods_DA_001", "frac_hoods_DA_005", "frac_hoods_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"

get_plot = function(var){
  p = ggplot(stat_combined_reduced , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = stat_combined_reduced[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim + order_k , ncol = 12) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DA nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


p = get_plot("frac_hoods_DA_01")
ggsave(filename = paste0(figures.dir, "da/split_all__fdr_01", ".png"), plot = p, width = 20, height = 6)

p = get_plot("frac_hoods_DA_005")
ggsave(filename = paste0(figures.dir, "da/split_all__fdr_005", ".png"), plot = p, width = 20, height = 6)

p = get_plot("frac_hoods_DA_001")
ggsave(filename = paste0(figures.dir, "da/split_all__fdr_001", ".png"), plot = p, width = 20, height = 6)


```

### Groupped by same sim

```{r heatmap-da-split-sim, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "frac_hoods_DA_001", "frac_hoods_DA_005", "frac_hoods_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"

stat_combined_reduced = as.data.frame(stat_combined_reduced %>% group_by(id_sim , reducedDim_name, frac_hoods_DA_001, frac_hoods_DA_005, frac_hoods_DA_01 , n_markers , n_not_markers , round_sim) %>% dplyr::summarise(frac_hoods_DA_001 = mean(frac_hoods_DA_001 , na.rm = T) , frac_hoods_DA_005 = mean(frac_hoods_DA_005 , na.rm = T) , frac_hoods_DA_01 = mean(frac_hoods_DA_01 , na.rm = T)) )



get_plot = function(var){
  p = ggplot(stat_combined_reduced , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = stat_combined_reduced[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DA nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


p = get_plot("frac_hoods_DA_01")
ggsave(filename = paste0(figures.dir, "da/groupped_sim__fdr_01", ".png"), plot = p, width = 8, height = 6)

p = get_plot("frac_hoods_DA_005")
ggsave(filename = paste0(figures.dir, "da/groupped_sim__fdr_005", ".png"), plot = p, width = 8, height = 6)

p = get_plot("frac_hoods_DA_001")
ggsave(filename = paste0(figures.dir, "da/groupped_sim__fdr_001", ".png"), plot = p, width = 8, height = 6)





```


### Only one order-k, split by sims

```{r heatmap-da-one-combo, message = FALSE}



stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "frac_hoods_DA_001", "frac_hoods_DA_005", "frac_hoods_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"

stat_combined_reduced = stat_combined_reduced[stat_combined_reduced$order_k == "1_100" , ]

get_plot = function(var){
  p = ggplot(stat_combined_reduced , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = stat_combined_reduced[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DA nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


p = get_plot("frac_hoods_DA_01")
ggsave(filename = paste0(figures.dir, "da/one_nhood__fdr_01", ".png"), plot = p, width = 8, height = 6)

p = get_plot("frac_hoods_DA_005")
ggsave(filename = paste0(figures.dir, "da/one_nhood__fdr_005", ".png"), plot = p, width = 8, height = 6)

p = get_plot("frac_hoods_DA_001")
ggsave(filename = paste0(figures.dir, "da/one_nhood__fdr_001", ".png"), plot = p, width = 8, height = 6)



```

### Groupped by eveyrhting

```{r heatmap-da-group-all, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "frac_hoods_DA_001", "frac_hoods_DA_005", "frac_hoods_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"

stat_combined_reduced = as.data.frame(stat_combined_reduced %>% group_by(reducedDim_name, frac_hoods_DA_001, frac_hoods_DA_005, frac_hoods_DA_01 , n_markers , n_not_markers ) %>% dplyr::summarise(frac_hoods_DA_001 = mean(frac_hoods_DA_001 , na.rm = T) , frac_hoods_DA_005 = mean(frac_hoods_DA_005 , na.rm = T) , frac_hoods_DA_01 = mean(frac_hoods_DA_01 , na.rm = T)) )



get_plot = function(var){
  p = ggplot(stat_combined_reduced , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = stat_combined_reduced[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DA nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes\ncontributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


p = get_plot("frac_hoods_DA_01")
ggsave(filename = paste0(figures.dir, "da/groupped_all__fdr_01", ".png"), plot = p, width = 5, height = 3)

p = get_plot("frac_hoods_DA_005")
ggsave(filename = paste0(figures.dir, "da/groupped_all__fdr_005", ".png"), plot = p, width = 5, height = 3)

p = get_plot("frac_hoods_DA_001")
ggsave(filename = paste0(figures.dir, "da/groupped_all__fdr_001", ".png"), plot = p, width = 5, height = 3)



```


## DE - overall

### Split by all

```{r heatmap-de-split-all, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_001", "frac_hoods_DE_005", "frac_hoods_DE_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"

get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim + order_k , ncol = 12) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de/" , as.character(anno$gene_type[i]), "/split_all__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 20, height = 6)
})





```


### Groupped by same sim

```{r heatmap-de-split-sim, message = FALSE}



stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_001", "frac_hoods_DE_005", "frac_hoods_DE_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"


stat_combined_reduced = as.data.frame(stat_combined_reduced %>% group_by(id_sim , reducedDim_name, gene_type , n_markers , n_not_markers , round_sim) %>% dplyr::summarise(frac_hoods_DE_001 = mean(frac_hoods_DE_001 , na.rm = T) , frac_hoods_DE_005 = mean(frac_hoods_DE_005 , na.rm = T) , frac_hoods_DE_01 = mean(frac_hoods_DE_01 , na.rm = T)) )



get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim ) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de/" , as.character(anno$gene_type[i]), "/groupped_sim__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 8, height = 6)
})



```


### Only one order-k, split by sims

```{r heatmap-de-one-combo, message = FALSE}



stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_001", "frac_hoods_DE_005", "frac_hoods_DE_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"


stat_combined_reduced = stat_combined_reduced[stat_combined_reduced$order_k == "1_100" , ]


get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim ) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de/" , as.character(anno$gene_type[i]), "/one_nhood__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 8, height = 6)
})


```



### Groupped by eveyrhting

```{r heatmap-de-group-all, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_001", "frac_hoods_DE_005", "frac_hoods_DE_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"


stat_combined_reduced = as.data.frame(stat_combined_reduced %>% group_by(reducedDim_name, gene_type , n_markers , n_not_markers , round_sim) %>% dplyr::summarise(frac_hoods_DE_001 = mean(frac_hoods_DE_001 , na.rm = T) , frac_hoods_DE_005 = mean(frac_hoods_DE_005 , na.rm = T) , frac_hoods_DE_01 = mean(frac_hoods_DE_01 , na.rm = T)) )



get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name ) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods" ) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes\ncontributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de/" , as.character(anno$gene_type[i]), "/groupped_all__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 5, height = 3)
})




```

## DE within not DA

### Split by all

```{r heatmap-de-not-da-split-all, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_not_DA_001", "frac_hoods_DE_not_DA_005", "frac_hoods_DE_not_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"

get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim + order_k , ncol = 12) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods\n(in not DA nhoods)") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_not_DA_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de_in_not_da/" , as.character(anno$gene_type[i]), "/split_all__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 20, height = 6)
})





```


### Groupped by same sim

```{r heatmap-de-not-da-split-sim, message = FALSE}



stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_not_DA_001", "frac_hoods_DE_not_DA_005", "frac_hoods_DE_not_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"


stat_combined_reduced = as.data.frame(stat_combined_reduced %>% group_by(id_sim , reducedDim_name, gene_type , n_markers , n_not_markers , round_sim) %>% dplyr::summarise(frac_hoods_DE_001 = mean(frac_hoods_DE_not_DA_001 , na.rm = T) , frac_hoods_DE_005 = mean(frac_hoods_DE_not_DA_005 , na.rm = T) , frac_hoods_DE_01 = mean(frac_hoods_DE_not_DA_01 , na.rm = T)) )



get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim ) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods\n(in not DA nhoods)") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de_in_not_da/" , as.character(anno$gene_type[i]), "/groupped_sim__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 8, height = 6)
})



```


### Only one order-k, split by sims

```{r heatmap-de-not-da-one-combo, message = FALSE}



stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_not_DA_001", "frac_hoods_DE_not_DA_005", "frac_hoods_DE_not_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"


stat_combined_reduced = stat_combined_reduced[stat_combined_reduced$order_k == "1_100" , ]


get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name + round_sim ) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods\n(in not DA nhoods)") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes contributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_not_DA_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de_in_not_da/" , as.character(anno$gene_type[i]), "/one_nhood__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 8, height = 6)
})


```


### Groupped by eveyrhting

```{r heatmap-de-not-da-group-all, message = FALSE}


stat_combined_reduced = unique(stat_combined[, c("id_sim" , "id_nhood_assignment" , "reducedDim_name", "gene_type", "frac_hoods_DE_not_DA_001", "frac_hoods_DE_not_DA_005", "frac_hoods_DE_not_DA_01" , "n_markers" , "n_not_markers" , "round_sim" , "order" , "k" , "seed")] )
stat_combined_reduced$order_k = paste0(stat_combined_reduced$order , "_" , stat_combined_reduced$k)

stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.hvgs"] = "Supervised"
stat_combined_reduced$reducedDim_name[stat_combined_reduced$reducedDim_name == "pca.all"] = "Unsupervised"


stat_combined_reduced = as.data.frame(stat_combined_reduced %>% group_by(reducedDim_name, gene_type , n_markers , n_not_markers , round_sim) %>% dplyr::summarise(frac_hoods_DE_001 = mean(frac_hoods_DE_not_DA_001 , na.rm = T) , frac_hoods_DE_005 = mean(frac_hoods_DE_not_DA_005 , na.rm = T) , frac_hoods_DE_01 = mean(frac_hoods_DE_not_DA_01 , na.rm = T)) )



get_plot = function(var , gene_type){
  current.stat = stat_combined_reduced[stat_combined_reduced$gene_type == gene_type , ]
  p = ggplot(current.stat , aes(x = factor(n_markers) , y = factor(n_not_markers) , fill = current.stat[,var])) +
    geom_tile() + 
    facet_wrap(~reducedDim_name ) + 
    scale_fill_viridis(discrete = F , name = "Fraction of DE nhoods" ) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    labs(x = "# of altered genes contributing to joint variance" , y = "# of altered genes\ncontributing to case variance") +
    theme(legend.position = "top")
  p
  return(p)
}


anno = expand.grid(gene_type = c("marker" , "not_marker") , var = c("01" , "005" , "001"))


plots = lapply(1:nrow(anno) , function(i){
  p = get_plot(paste0("frac_hoods_DE_" ,as.character(anno$var[i])) , as.character(anno$gene_type[i]))
  ggsave(filename = paste0(figures.dir, "de_in_not_da/" , as.character(anno$gene_type[i]), "/groupped_all__fdr_" , as.character(anno$var[i]), ".png"), plot = p, width = 5, height = 3)
})


```

## DA vs DE

@@ Mb later, lets see if we need it

```{r da-vs-de, message = FALSE}




```

# UMAP cartoons

## Plotting function

```{r plotting-function, message = FALSE}


cols = met.brewer("Egypt",4)
cols = cols[c(3:4)]

get_plot = function(id , red_dim){
  sce = sces[[id]]

  umaps = as.data.frame( scater::calculateUMAP(t(reducedDim(sce , red_dim))) )
  reducedDim(sce , "umap") = umaps
  meta = cbind(as.data.frame(colData(sce)) , as.data.frame(reducedDim(sce , "umap")))
  meta$Group = as.character(meta$Group)
  meta$Group[meta$Group == "Group1"] = "Cell type 1"
  meta$Group[meta$Group == "Group2"] = "Cell type 2"
  meta$type[meta$type == "cond_A"] = "Condition A"
  meta$type[meta$type == "cond_B"] = "Condition B"
  
  meta = meta[sample(nrow(meta)) , ]
  #title = names(sces)[id]
  #title = str_remove(title , "__n_not_markers_0")
  p = ggplot(meta , aes(x = V1 , y = V2 , col = factor(Group))) + 
    geom_point(alpha = .5 , size = .5) +
    theme_bw() +
    facet_wrap(~type, nrow = 2) + 
    labs(x = "UMAP-1" , y = "UMAP-2") + 
    scale_color_manual(values = cols , name = "Cell type") 
  p
  return(p)
}





```

## Base simulation

```{r umap-base, message = FALSE}


id = 1
p = get_plot(id, "pca.hvgs")
p
ggsave(filename = paste0(figures.dir, "umaps/hvgs/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)


p = get_plot(id, "pca.all")
p
ggsave(filename = paste0(figures.dir, "umaps/all/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)



```

## Only marker perturbed

```{r umap-marker, message = FALSE}


id = 25
p = get_plot(id, "pca.hvgs")
p
ggsave(filename = paste0(figures.dir, "umaps/hvgs/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)


p = get_plot(id, "pca.all")
p
ggsave(filename = paste0(figures.dir, "umaps/all/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)



```

## Only not markers perturbed

```{r umap-not-marker, message = FALSE}


id = 361
p = get_plot(id, "pca.hvgs")
p
ggsave(filename = paste0(figures.dir, "umaps/hvgs/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)


p = get_plot(id, "pca.all")
p
ggsave(filename = paste0(figures.dir, "umaps/all/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)



```

## Both perturbed

```{r umaps-both, message = FALSE}


id = 385
p = get_plot(id, "pca.hvgs")
p
ggsave(filename = paste0(figures.dir, "umaps/hvgs/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)


p = get_plot(id, "pca.all")
p
ggsave(filename = paste0(figures.dir, "umaps/all/" , names(sces)[id], ".png"), plot = p, width = 4, height = 5)



```


## Overall grid (split by sims)


```{r umaps-all, message = FALSE}



get_all_plots = function(round_sim , red_dim){
  ids = sapply(names(sces) , function(str) str_detect(str , paste0(round_sim, "_n_markers_"), negate = FALSE))
  ids = as.numeric( which(ids) )
  
  plots = bplapply(ids , function(id){
    print(id)
    p = get_plot(id , red_dim)
    return(p)
  }, BPPARAM = mcparam)
  p = ggarrange(plotlist = plots , nrow = 12 , ncol = 12 , common.legend = T)
  return(p)
}


anno = expand.grid(round_sim = c(1,2,3) , red_dim = c("all" , "hvgs"))

plots = lapply(1:nrow(anno) , function(i){
  p = get_all_plots(as.numeric(as.character(anno$round_sim[i])) , red_dim = paste0("pca.", as.character(anno$red_dim[i])))
  ggsave(filename = paste0(figures.dir, "umaps/" , as.character(anno$red_dim[i]), "/" , "all_" , as.character(anno$round_sim[i]), ".png"), plot = p, width = 17, height = 22)
})




```



# Session Info

```{r sessinf}
sessionInfo()
```
