---
title: "This pipeline to generate chimera dataset ( WT + chimera, different embeddings)"
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---


# Load dependencies

```{r load, message = FALSE}


library(SingleCellExperiment)
library(MouseGastrulationData)
library(scran)
library(uwot)
library(BiocParallel)
library(zellkonverter)
ncores = 2
mcparam = MulticoreParam(workers = ncores)
register(mcparam)

root.dir = "/nfs/research/marioni/alsu/hubmap_metaRef/"
source(paste0(root.dir , "miloDE_analysis/core_functions.R"))

set.seed(32)


```

# Load WT data (E8.5)

```{r load-wt-data, message = FALSE}


sce_ref = readRDS(paste0(root.dir , "data/sces/mouse_embryo/WT/mouse_embryo_E8.5.Rds"))
sce_ref$type = "wt"
sce_ref$tal1 = 1
  

```

# Load chimera data (Tal1 one)

```{r load-tal-data, message = FALSE}


sce_chimera = Tal1ChimeraData()

# select tomato positive cells (tal1 KO); delete doublets and stripped
sce_chimera = sce_chimera[, !sce_chimera$celltype.mapped %in% c("Doublet" , "Stripped")]

# delete row for tomato
sce_chimera = sce_chimera[!rownames(sce_chimera) == "tomato-td" , ]

# add chimera_ prefix to avoid conclusion
colnames(sce_chimera) = paste0("chimera_" , colnames(sce_chimera) )
sce_chimera$cell = paste0("chimera_" ,sce_chimera$cell)
sce_chimera$sample = paste0("chimera_" , sce_chimera$sample )

sce_chimera$type = "chimera"
sce_chimera$tal1 = ifelse(sce_chimera$tomato , 0 , 1)


```

# Concatenate and add logcounts

```{r add-logcounts, message = FALSE}


# order rownames, check that dss are concatenable
sce_ref = sce_ref[order(rownames(sce_ref)) , ]
sce_chimera = sce_chimera[order(rownames(sce_chimera)) , ]
print(mean(rownames(sce_ref) == rownames(sce_chimera)))

# concatenate
sce = SingleCellExperiment(list(counts = cbind(counts(sce_ref) , counts(sce_chimera))))
# add metadata
sce$cell = c(sce_ref$cell , sce_chimera$cell)
sce$celltype = c(sce_ref$celltype , sce_chimera$celltype.mapped)
sce$sample = c(sce_ref$sample , sce_chimera$sample)
sce$type = c(sce_ref$type , sce_chimera$type)
sce$tal1 = c(sce_ref$tal1 , sce_chimera$tal1)

# add rowdata
rowData(sce) = rowData(sce_ref)

# add logcounts
sce = add_logcounts(sce)


```

# Add reduced dims (embeddings)

## MNN-corrected PCs (both modalities)

### HVGs - ref + query

```{r mnn-pcs-hvgs-all, message = FALSE}


hvgs <- getTopHVGs(sce, n=3000)
sce = add_batch_corrected_pca(sce , hvgs , reduced.dim_name = "pca.corrected_all" , bpparam = mcparam)

# add umaps
sce = scater::runUMAP(sce , dimred = "pca.corrected_all" , name = "UMAP_pca.corrected_all" , ncomponents = 2)


```

### HVGs - ref only

```{r mnn-pcs-hvgs-ref, message = FALSE}


hvgs <- getTopHVGs(sce[, sce$type == "wt"], n=3000)
sce = add_batch_corrected_pca(sce , hvgs , reduced.dim_name = "pca.corrected_wt" , bpparam = mcparam)

# add umaps
sce = scater::runUMAP(sce , dimred = "pca.corrected_wt" , name = "UMAP_pca.corrected_wt" , ncomponents = 2)


```

### HVGs - query only

```{r mnn-pcs-hvgs-query, message = FALSE}


hvgs <- getTopHVGs(sce[, !sce$type == "wt"], n=3000)
sce = add_batch_corrected_pca(sce , hvgs , reduced.dim_name = "pca.corrected_chimera" , bpparam = mcparam)

# add umaps
sce = scater::runUMAP(sce , dimred = "pca.corrected_chimera" , name = "UMAP_pca.corrected_chimera" , ncomponents = 2)


```

## MNN-corrected PCs (only ref)

```{r mnn-only-ref, message = FALSE}


sce_wt = sce[, sce$type == "wt"]
sce_chimera = sce[, !sce$type == "wt"]
hvgs <- getTopHVGs(sce_wt, n=3000)


full_pca = multiBatchPCA(sce_wt , batch = factor(sce_wt$sample) , subset.row = hvgs, d = 30, preserve.single = TRUE, assay.type = "logcounts")
ref_pca <- full_pca[[1]]
to_proj <- (as.matrix(logcounts(sce_chimera)[hvgs,]) - rowMeans(as.matrix(logcounts(sce_chimera)[hvgs,])))/sqrt(rowVars(as.matrix(logcounts(sce_chimera)[hvgs,])))

map_pca <- t(to_proj) %*% metadata(full_pca)$rotation
big_pca <- rbind(ref_pca, map_pca)
# check that rows are ordered correctly
mean(rownames(big_pca) == colnames(sce))

# correct
out = reducedMNN(big_pca , batch = factor(sce$sample) )
joint.pca = out$corrected
reducedDim(sce , "pca.corrected_ref.proj") = joint.pca

# add umaps
sce = scater::runUMAP(sce , dimred = "pca.corrected_ref.proj" , name = "UMAP_pca.corrected_ref.proj" , ncomponents = 2)



```


## Azimuth

```{r azimuth, message = FALSE}


sce = add_azimuth_pca(sce , ref_samples = unique(sce$sample[sce$type == "wt"]), query_samples = unique(sce$sample[!sce$type == "wt"]), nPC = 30 , bpparam = mcparam , 
                      reducedDim.name = "azimuth" )
# add umaps
sce = scater::runUMAP(sce , dimred = "azimuth" , name = "UMAP_azimuth" , ncomponents = 2)

saveRDS(sce , file = paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal.Rds"))


```


## scVI

Download results from python script here

```{r add scvi, message = FALSE}


sce = readRDS(file = paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal.Rds"))

sce_scvi = readH5AD(paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal_w_scarches.h5ad") , reader = "R")
# correct colnames
colnames(sce_scvi) = sce_scvi$cell

# reorder sce and sce-azimuth
sce = sce[ , order(colnames(sce))]
sce_scvi = sce_scvi[ , order(colnames(sce_scvi))]

print(mean(colnames(sce) == colnames(sce_scvi)))

reducedDim(sce , "scVI") = reducedDim(sce_scvi , "X_scVI")

# add umaps
sce = scater::runUMAP(sce , dimred = "scVI" , name = "UMAP_scVI" , ncomponents = 2)

saveRDS(sce , file = paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal.Rds"))


```


# Add subclusters

## Brain

### Add louvain clustering

```{r add-subclusters-brain-get-louvain, message = FALSE}


sce = readRDS(file = paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal.Rds"))

sce$sub_celltype = sce$celltype

ct = "Forebrain/Midbrain/Hindbrain"
sce_ct = sce[, sce$celltype == ct]
reducedDim(sce_ct , "UMAP_scVI") = as.data.frame(uwot::umap(reducedDim(sce_ct , "scVI") , min_dist = 1))

sce_ct = retain_informative_genes(sce_ct , n = 1000) 
snn_graph = buildSNNGraph(reducedDim(sce_ct , "scVI"), k = 50 , d = 10 , transposed = T)
clusters = cluster_louvain(snn_graph)  
sce_ct$cluster = clusters$membership
sce_ct$cluster = as.factor(sce_ct$cluster)

umaps = cbind(as.data.frame(reducedDim(sce_ct , "UMAP_scVI")) , as.data.frame(colData(sce_ct)))
p = ggplot(umaps , aes(x = V1 , y = V2 , col = cluster)) + 
  geom_point() + 
  theme_bw() + 
  theme(legend.position = "top")
p


```

### Plot UMAPs + counts

```{r check-how-markers-look, message = FALSE}



rowdata = as.data.frame(rowData(sce_ct))
get_plot_w_colors = function(sce , gene){
  umaps = cbind(as.data.frame(reducedDim(sce , "UMAP_scVI")) , as.data.frame(colData(sce)))
  ensembl = rowdata$ENSEMBL[rowdata$SYMBOL == gene]
  umaps$counts = as.numeric(logcounts(sce[ensembl , ]))
  umaps = umaps[order(umaps$counts) , ]
  p = ggplot(umaps , aes(x = V1 , y = V2 , col = counts)) +
    geom_point(size=.75) + 
    scale_color_viridis(discrete = F) +
    theme_bw() + 
    ggtitle(gene) + 
    theme(legend.position = "none")
  return(p)
}


genes = c("Shh" , "Rax" , "Six3" , "Otx2" , "En1" , "Hoxb2" , "Hoxa2" , "Gbx2")
plots = lapply(genes , function(gene){
  p = get_plot_w_colors(sce_ct , gene)
  return(p)
})
p = ggarrange(plotlist = plots , ncol = 4 , nrow = 2)
p




```

Clusters:

7 -- floor plate
4,5 -- forebrain
1,6 -- midbrain
2,3 -- hindbrain

### Assign normal identities for clusters

```{r add-subclusters-brain, message = FALSE}


clusters = data.frame(cluster = c(1:7) , cluster_name = c("Midbrain" , "Hindbrain" , "Hindbrain" , "Forebrain" , "Forebrain" , "Midbrain" , "Floor plate"))

sce_ct$sub_celltype = sapply(sce_ct$cluster , function(x) clusters$cluster_name[clusters$cluster == x])
sce_ct$sub_celltype = factor(sce_ct$sub_celltype , sort(unique(sce_ct$sub_celltype)))

# very ugly code but v quick to write
for (sub_ct in unique(sce_ct$sub_celltype)){
  print(sub_ct)
  cells = sce_ct$cell[sce_ct$sub_celltype == sub_ct]
  idx = which(sce$cell %in% cells)
  sce$sub_celltype[idx] = sub_ct
}
saveRDS(sce , file = paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal.Rds"))



```



# Session Info

```{r sessinf}
sessionInfo()
```
