---
title: "This pipeline to test different embedding strategies"
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---
# Load dependencies

```{r load, message = FALSE}


library(SingleCellExperiment)
library(miloDE)
library(miloR)
library(BiocParallel)
library(ggplot2)
library(ggpubr)
library(geneBasisR)
library(MouseGastrulationData)
library(scran)
library(MetBrewer)

ncores = 7
mcparam = MulticoreParam(workers = ncores)
register(mcparam)
set.seed(32)


root.dir = "/nfs/research/marioni/alsu/hubmap_metaRef/"
figures.dir = paste0(root.dir , "figures/optimal_coverage/embedding_updated/")


# load sce with embedding
sce = readRDS(paste0(root.dir , "data/sces/mouse_embryo/chimera_WT/mouse_embryo_wt_chimera.Rds"))
rowdata = as.data.frame(rowData(sce))

# load genes
genes_wt = read.csv(paste0(root.dir , "data/sces/mouse_embryo/chimera_WT/HVGs/wt.csv") , row.names = 1)
genes_all = read.csv(paste0(root.dir , "data/sces/mouse_embryo/chimera_WT/HVGs/all.csv") , row.names = 1)
genes = list(all = genes_all$HVG , wt = genes_wt$HVG)


chimera_speicific_genes = c("Rpgrip1" , "Cbx3" , "Actb")
chimera_speicific_stat = data.frame(SYMBOL = chimera_speicific_genes)
chimera_speicific_stat = merge(chimera_speicific_stat , rowdata , all.x = T , all.y = F)
chimera_speicific_stat$gene_in_wt = chimera_speicific_stat$ENSEMBL %in% genes$wt
chimera_speicific_stat$gene_in_all = chimera_speicific_stat$ENSEMBL %in% genes$all
chimera_speicific_stat$SYMBOL = factor(chimera_speicific_stat$SYMBOL , levels = chimera_speicific_genes)


anno_mnn = expand.grid(type = c("supervised" , "unsupervised") , genes = c("wt", "all") , method = c("mnn") )
anno_azimuth = expand.grid(type = c("supervised" , "unsupervised") , genes = c("wt", "all") , method = c("azimuth" ))
anno_rest = expand.grid(type = c("supervised" , "unsupervised") , genes = c("wt", "all") , method = c( "scvi" , "scanvi"))
anno = rbind(anno_azimuth , anno_rest , anno_mnn)
anno$reducedDim.name = paste0(anno$method , "_" , anno$type , "_" , anno$genes)


sce_milo_nhoods = readRDS(paste0(root.dir , "data/processed/embedding_analysis_updated/sce_milo_nhoods.Rds"))

stat = readRDS(paste0(root.dir , "data/temp/embedding_homogeneity_stat.Rds"))


```


# UMAPs, colored by CT

@@ Old, dont need it probably

```{r umaps-by-ct, message = FALSE}



get_umap_plot = function(sce , reducedDim.name){
  set.seed(32)
  umaps = as.data.frame( scater::calculateUMAP(t(reducedDim(sce , reducedDim.name))) )
  umaps = cbind(umaps , as.data.frame(colData(sce)))
  umaps$type = factor(umaps$type , levels = c("wt" , "chimera"))
  p = ggplot(umaps , aes(x = V1 , y = V2 , col = celltype)) + 
    geom_point(size=0.5) + 
    theme_bw() + 
    theme(legend.position = "none") + 
    scale_color_manual(values = EmbryoCelltypeColours) + 
    ggtitle(reducedDim.name) +
    facet_wrap(~type , nrow = 2) +
    labs(x = "" , y = "") 
  return(p)
}

plots = lapply(1:nrow(anno), function(i){
  reducedDim.name = anno$reducedDim.name[i]
  p = get_umap_plot(sce , reducedDim.name)
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = 4 , nrow = 4)
p
ggsave(filename = paste0(figures.dir, "umaps_by_ct", ".png"), plot = p, width = 10, height = 15)



```

@@ Add mapping

# Select chimera-specific genes for testing

```{r check-chimera-genes, message = FALSE}



get_de_genes_bulk = function(sce , celltype = NULL){
  require(tibble)
  require(edgeR)
  if (!is.null(celltype)){
    sce = sce[ , sce$celltype == celltype]
  }
  
  stat = tryCatch(
    {
      summed = summarizeAssayByGroup(counts(sce), colData(sce)[,c("sample","type")])
      y <- DGEList(assay(summed, "sum"), samples=colData(summed), lib.size = colSums(assay(summed , "sum")))
      keep <- filterByExpr(y, group=summed$sample , min.count = 2 , min.total.count = 4)
      y <- y[keep,]
      y <- calcNormFactors(y)
      design <- model.matrix(~ factor(type), y$samples)
      y <- estimateDisp(y, design)
      fit <- glmQLFit(y, design, robust=TRUE)
      res <- glmQLFTest(fit, coef=ncol(design))
      out = topTags(res, n = Inf )
      out = as.data.frame(out)
      out = rownames_to_column(out , var = "ENSEMBL")
      out = merge(out , rowdata , all.x = T , all.y = F)
      if (!is.null(celltype)){
        out$celltype = celltype
      }
      out
    }, 
    error = function(dummy){
      return(NULL)
    }
  )
  return(stat)
}      



de_stat_bulk = lapply(unique(sce$celltype) , function(celltype){
  print(celltype)
  out = get_de_genes_bulk(sce , celltype)
  return(out)
})
de_stat_bulk = do.call(rbind , de_stat_bulk)

# get celltype composition
tab = as.data.frame(table(sce$celltype))
colnames(tab) = c("celltype" , "n")
tab$n = tab$n/sum(tab$n)

de_stat_bulk = merge(de_stat_bulk , tab , all.x = T , all.y = F)
# add normalised celltype composition
de_stat_bulk$n_norm = de_stat_bulk$n
de_stat_bulk$n_norm[de_stat_bulk$PValue >= 0.01] = 0
   

de_stat_per_gene = as.data.frame(de_stat_bulk %>% group_by(ENSEMBL , SYMBOL) %>% dplyr::summarise(median_pval = median(PValue , na.rm = T) , median_logFC = median(logFC , na.rm = T) , mean_logFC = mean(logFC , na.rm = T) , frac_cts_sig = mean(FDR < 0.1 , na.rm = T) , 
                                                                                                  weight_cts_sig = sum(n_norm)))
de_stat_per_gene = de_stat_per_gene[de_stat_per_gene$SYMBOL %in% as.character(chimera_speicific_stat$SYMBOL) , ]


# add DE using all
de_stat_all = get_de_genes_bulk(sce)
de_stat_all = de_stat_all[de_stat_all$ENSEMBL %in% chimera_speicific_stat$ENSEMBL , c("ENSEMBL" , "logFC")]
colnames(de_stat_all) = c("ENSEMBL" , "logFC_all")

de_stat_per_gene = merge(de_stat_per_gene , de_stat_all)


```

## Boxplot w/ counts

```{r boxplot-w-chimera-specific-counts, message = FALSE}


cols = wesanderson::wes_palette("Royal1" , 2)
names(cols) = c("wt" , "chimera")
get_boxplot_per_gene = function(gene){
  df = data.frame(cell = sce$cell , sample = sce$sample , type = sce$type , 
                  counts = as.numeric(logcounts(sce[rowdata$ENSEMBL[rowdata$SYMBOL == gene] , ])))
  df = df[order(df$counts) , ]
  df$type = factor(df$type , levels = c("wt" , "chimera"))
  p = ggplot(df , aes(x = type , y = counts , fill = type)) + 
    geom_boxplot(alpha = .9) + 
    theme_bw() + 
    theme(legend.position = "right") + 
    scale_fill_manual(values = cols , name = "Condition") +
    ggtitle(gene ) +
    theme(axis.text.x = element_blank()) +
    labs(x = "" , y = "logcounts") 
    #theme(legend.text=element_text(size=15) , legend.title=element_text(size=20))
  return(p)
}



plots = lapply(levels(chimera_speicific_stat$SYMBOL) , function(gene){
  p = get_boxplot_per_gene(gene) +
    theme(legend.position = "none")
  print(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(chimera_speicific_stat$SYMBOL))
p
ggsave(filename = paste0(figures.dir, "boxplot_w_chimera_genes_counts", ".png"), plot = p, width = 6, height = 2)



```


# Neighbourhood homogeneity

## Get homogeneity stat

```{r stat-heterogeneity-hoods, message = FALSE}



get_mean_sd_distr_per_hood = function(idx_milo , gene){
  df = data.frame(cell = colnames(sce) , 
                            sample = sce$sample,
                            type = sce$type , 
                            counts = as.numeric(logcounts(sce[rowdata$ENSEMBL[rowdata$SYMBOL == as.character(gene)] , ])))
  nhoods_sce = sce_milo_nhoods[[idx_milo]]
  stat = lapply(1:ncol(nhoods_sce) , function(i){
    idx = which(nhoods_sce[,i] == 1 )
    current.df = df[df$cell %in% rownames(nhoods_sce)[idx],]
    current.stat = as.data.frame(current.df %>% group_by(type) %>% dplyr::summarise(mean = mean(counts) , sd = sd(counts)))
    current.stat$Nhood = i
    return(current.stat)
  })
  stat = do.call(rbind , stat)
  id = names(sce_milo_nhoods)[idx_milo]
  stat$id = id
  id = stringr::str_split(id , pattern = "_")
  id = id[[1]]
  stat$reduced_dim_method = id[1]
  stat$reduced_dim_type = id[2]
  stat$reduced_dim_genes = id[3]
  stat$reduced_dim = paste0(id[1] , "_" , id[2] , "_" , id[3])
  stat$order = as.numeric(id[4])
  stat$k = as.numeric(id[5])
  stat$round = as.numeric(id[6])
  return(stat)
}

anno = expand.grid(gene = chimera_speicific_stat$SYMBOL , idx_milo = c(1:length(sce_milo_nhoods)))
stat = bplapply(1:nrow(anno) , function(i){
  out = get_mean_sd_distr_per_hood(anno$idx_milo[i] , as.character(anno$gene[i]))
  out$gene = as.character(anno$gene[i])
  return(out)
} , BPPARAM = mcparam)
stat = do.call(rbind , stat)
stat = stat[order(stat$mean) , ]
stat$type = factor(stat$type , levels = c("wt" , "chimera"))
saveRDS(stat , paste0(root.dir , "data/temp/embedding_homogeneity_stat.Rds"))



```


# Plot

## Single simulation

```{r plot-sd-boxplot, message = FALSE}


anno_mnn = expand.grid(type = c("supervised" , "unsupervised") , genes = c("wt", "all") , method = c("mnn") )
anno_azimuth = expand.grid(type = c("supervised" , "unsupervised") , genes = c("wt", "all") , method = c("azimuth" ))
anno_rest = expand.grid(type = c("supervised" , "unsupervised") , genes = c("wt", "all") , method = c( "scvi" , "scanvi"))
anno = rbind(anno_azimuth , anno_rest , anno_mnn)
anno$reducedDim.name = paste0(anno$method , "_" , anno$type , "_" , anno$genes)


stat = readRDS(paste0(root.dir , "data/temp/embedding_homogeneity_stat.Rds"))
# fix for one round
stat = stat[stat$round == 1 , ]

current.stat = stat[stat$reduced_dim %in% anno$reducedDim.name,]
current.stat$reduced_dim = factor(current.stat$reduced_dim , levels = anno$reducedDim.name)
current.stat$reduced_dim_type_genes = paste0(current.stat$reduced_dim_type , "_" , current.stat$reduced_dim_genes)
current.stat$reduced_dim_type_genes = factor(current.stat$reduced_dim_type_genes , levels = c("unsupervised_all" , "unsupervised_wt" , "supervised_all" , "supervised_wt"))

# update naming
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "azimuth"] = "Azimuth"
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "mnn"] = "MNN"
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "scvi"] = "scVI"
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "scanvi"] = "scANVI"

#cols = c( "paleturquoise3" , "paleturquoise4" , "rosybrown2" , "rosybrown4" )
cols = met.brewer("Derain" , n = 7)[c(1,2,4,5)]


plots = lapply(levels(chimera_speicific_stat$SYMBOL) , function(gene){
  current.x = current.stat$sd[current.stat$gene == gene & current.stat$type == "chimera"]
  current.x = current.x[!is.na(current.x)]
  p = ggplot(current.stat[current.stat$gene == gene & current.stat$type == "chimera", ] , aes(x = reduced_dim_method , y = sd , fill = reduced_dim_type_genes)) + 
    geom_boxplot(outlier.size= .25 , fatten = 1 , lwd = .2) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    labs(x = "" , y = "")  +
    scale_fill_manual(values = cols, name = "Embedding") +
    theme(legend.position = "none") + ggtitle(gene) +
    ylim(c(quantile(current.x , 0.0) , quantile(current.x , 1) ))
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(levels(chimera_speicific_stat$SYMBOL)) , nrow = 1)
p
ggsave(filename = paste0(figures.dir, "sd_distr", ".png"), plot = p, width = 6.5, height = 2.8)


# assess significance (within same method)
anno_sig.grid = expand.grid(gene = as.character(unique(chimera_speicific_stat$SYMBOL)) , method = as.character(unique(current.stat$reduced_dim_method)) , reduced_dim_type_genes = c("supervised_wt" , "supervised_all" , "unsupervised_wt"))

sig_stat = lapply(1:nrow(anno_sig.grid) , function(i){
  current.gene = anno_sig.grid$gene[i]
  current.method = anno_sig.grid$method[i]
  
  x = current.stat$sd[current.stat$gene == current.gene & current.stat$reduced_dim_method == current.method & current.stat$reduced_dim_type_genes == "unsupervised_all"]
  y = current.stat$sd[current.stat$gene == current.gene & current.stat$reduced_dim_method == current.method & current.stat$reduced_dim_type_genes == as.character(anno_sig.grid$reduced_dim_type_genes[i])]
  x = x[!is.na(x)]
  y = y[!is.na(y)]
  
  out = t.test(x , y)  
  out = out$p.value
  out = data.frame(gene = current.gene , method = current.method , reduced_dim_type_genes = as.character(anno_sig.grid$reduced_dim_type_genes[i]) , pval = out)
  return(out)
})
sig_stat = do.call(rbind , sig_stat)
sig_stat$pval_005 = sig_stat$pval < 0.05
sig_stat$pval_001 = sig_stat$pval < 0.01
sig_stat$pval_0001 = sig_stat$pval < 0.001
sig_stat = sig_stat[!sig_stat$method == "Azimuth" , ]



```

## Combined across several simulations

```{r plot-sd-boxplot, message = FALSE}


stat = readRDS(paste0(root.dir , "data/temp/embedding_homogeneity_stat.Rds"))

stat = as.data.frame(stat[stat$type == "chimera", ] %>% group_by(reduced_dim_method , reduced_dim_type , reduced_dim_genes , reduced_dim , gene , round) %>% dplyr::summarise(sd = mean(sd , na.rm = T)))


current.stat = stat[stat$reduced_dim %in% anno$reducedDim.name,]
current.stat$reduced_dim = factor(current.stat$reduced_dim , levels = anno$reducedDim.name)
current.stat$reduced_dim_type_genes = paste0(current.stat$reduced_dim_type , "_" , current.stat$reduced_dim_genes)
current.stat$reduced_dim_type_genes = factor(current.stat$reduced_dim_type_genes , levels = c("unsupervised_all" , "unsupervised_wt" , "supervised_all" , "supervised_wt"))

# update naming
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "azimuth"] = "Azimuth"
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "mnn"] = "MNN"
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "scvi"] = "scVI"
current.stat$reduced_dim_method[current.stat$reduced_dim_method == "scanvi"] = "scANVI"

plots = lapply(levels(chimera_speicific_stat$SYMBOL) , function(gene){
  p = ggplot(current.stat[current.stat$gene == gene , ] , aes(x = reduced_dim_method , y = sd , fill = reduced_dim_type_genes)) + 
    geom_boxplot(outlier.size= .25 , fatten = 1 , lwd = .2) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    labs(x = "" , y = "")  +
    scale_fill_manual(values = cols, name = "Embedding") +
    theme(legend.position = "none") + ggtitle(gene) 
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(levels(chimera_speicific_stat$SYMBOL)) , nrow = 1)
p
ggsave(filename = paste0(figures.dir, "sd_distr_combined", ".png"), plot = p, width = 6.5, height = 2.8)


```

# DE-stat

## load DE stat for select genes


```{r boxplot-distr-milo-de, message = FALSE}


de_stat = readRDS( paste0(root.dir , "data/processed/embedding_analysis_updated/milo_de_chimera_genes.Rds") )
rowdata = as.data.frame(rowData(sce))
colnames(de_stat)[colnames(de_stat) == "gene"] = "ENSEMBL"
de_stat = merge(de_stat , rowdata , all.x = T , all.y = F)
de_stat = de_stat[de_stat$SYMBOL %in% chimera_speicific_stat$SYMBOL , ]
de_stat$SYMBOL = factor(de_stat$SYMBOL , levels = chimera_speicific_genes)
de_stat = merge(de_stat , de_stat_per_gene[, c("ENSEMBL" , "median_logFC" , "mean_logFC" , "logFC_all")] , all.x = T , all.y = F)

#
de_stat = de_stat[de_stat$reduced_dim %in% anno$reducedDim.name , ]
de_stat$reduced_dim = factor(de_stat$reduced_dim , levels = anno$reducedDim.name)
de_stat$reduced_dim_type_genes = paste0(de_stat$reduced_dim_type , "_" , de_stat$reduced_dim_genes)
de_stat$reduced_dim_type_genes = factor(de_stat$reduced_dim_type_genes , levels = c("unsupervised_all" , "unsupervised_wt" , "supervised_all" , "supervised_wt"))


# update methods
de_stat$reduced_dim_method[de_stat$reduced_dim_method == "azimuth"] = "Azimuth"
de_stat$reduced_dim_method[de_stat$reduced_dim_method == "mnn"] = "MNN"
de_stat$reduced_dim_method[de_stat$reduced_dim_method == "scvi"] = "scVI"
de_stat$reduced_dim_method[de_stat$reduced_dim_method == "scanvi"] = "scANVI"



```

## miloDE stat - single simulation

### logFC

```{r boxplot-distr-milo-de, message = FALSE}


current.de_stat = de_stat[de_stat$round == 1 , ]

plots = lapply(levels(de_stat$SYMBOL) , function(gene){
  current.x = abs(current.de_stat$logFC[current.de_stat$SYMBOL == gene])
  current.x = current.x[!is.na(current.x)]
  if (gene == "Cbx3"){
    ylim = c(0,5)
  } else if (gene == "Actb"){
    ylim = c(0,3)
  } else {
    ylim = c(2,10.5)
  }
  p = ggplot(current.de_stat[current.de_stat$SYMBOL == gene, ] , aes(x = reduced_dim_method , y = abs(logFC) , fill = reduced_dim_type_genes)) +
    geom_boxplot(outlier.size = .25, fatten = 1 , lwd = .2) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    ggtitle(gene) + 
    scale_fill_manual(values = cols , name = "Embedding") + 
    labs(x = "" , y = "") +
    geom_hline(yintercept = abs(unique(current.de_stat$logFC_all[current.de_stat$SYMBOL == gene])) , "color" = "red" , "linetype" = "dashed") +
    theme(legend.position = "none") +
    ylim(ylim)
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(chimera_speicific_genes) , nrow = 1)
p
ggsave(filename = paste0(figures.dir, "logFC_distr", ".png"), plot = p, width = 6.5, height = 2.8)


# stat
anno_sig.grid = expand.grid(gene = as.character(unique(chimera_speicific_stat$SYMBOL)) , method = as.character(unique(current.de_stat$reduced_dim_method)) , reduced_dim_type_genes = c("supervised_wt" , "supervised_all" , "unsupervised_wt"))

sig_stat = lapply(1:nrow(anno_sig.grid) , function(i){
  current.gene = anno_sig.grid$gene[i]
  current.method = anno_sig.grid$method[i]
  
  x = abs( current.de_stat$logFC[current.de_stat$SYMBOL == current.gene & current.de_stat$reduced_dim_method == current.method & current.de_stat$reduced_dim_type_genes == "unsupervised_all"] )
  y = abs( current.de_stat$logFC[current.de_stat$SYMBOL == current.gene & current.de_stat$reduced_dim_method == current.method & current.de_stat$reduced_dim_type_genes == as.character(anno_sig.grid$reduced_dim_type_genes[i])])
  x = x[!is.na(x)]
  y = y[!is.na(y)]
  
  out = t.test(x , y)  
  out = out$p.value
  out = data.frame(gene = current.gene , method = current.method , reduced_dim_type_genes = as.character(anno_sig.grid$reduced_dim_type_genes[i]) , pval = out)
  return(out)
})
sig_stat = do.call(rbind , sig_stat)
sig_stat$pval_005 = sig_stat$pval < 0.05
sig_stat$pval_001 = sig_stat$pval < 0.01
sig_stat$pval_0001 = sig_stat$pval < 0.001
sig_stat = sig_stat[!sig_stat$method == "Azimuth" , ]



```

### Frac of hood sig

```{r boxplot-distr-milo-de, message = FALSE}


de_stat$pval_corrected_across_nhoods[is.na(de_stat$pval_corrected_across_nhoods)] = 1
current.de_stat = de_stat[de_stat$round == 1 , ]

current.de_stat = as.data.frame(current.de_stat %>% group_by(SYMBOL , id , reduced_dim_method , reduced_dim_type , reduced_dim_genes , reduced_dim_type_genes , reduced_dim , round) %>% dplyr::summarise(frac_hoods_sig_01 = mean(pval_corrected_across_nhoods < 0.1 , na.rm = T) , frac_hoods_sig_005 = mean(pval_corrected_across_nhoods < 0.05 , na.rm = T) , frac_hoods_sig_001 = mean(pval_corrected_across_nhoods < 0.01 , na.rm = T) , frac_hoods_sig_0001 = mean(pval_corrected_across_nhoods < 0.005 , na.rm = T)  , frac_hoods_sig_0005 = mean(pval_corrected_across_nhoods < 0.001 , na.rm = T) , frac_hoods_sig_00001 = mean(pval_corrected_across_nhoods < 0.0001 , na.rm = T) ))


plots = lapply(levels(chimera_speicific_stat$SYMBOL) , function(gene){
  p = ggplot(current.de_stat[current.de_stat$SYMBOL == gene , ] , aes(x = reduced_dim_method , y = frac_hoods_sig_01 , fill = reduced_dim_type_genes)) + 
    geom_bar(stat = "identity" , position = "dodge") + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    labs(x = "" , y = "")  +
    scale_fill_manual(values = cols, name = "Embedding") +
    theme(legend.position = "none") + ggtitle(gene) 
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(levels(chimera_speicific_stat$SYMBOL)) , nrow = 1)
p
ggsave(filename = paste0(figures.dir, "frac_sig_distr", ".png"), plot = p, width = 6.5, height = 2.8)


```



## Combined

### logFC

```{r boxplot-distr-milo-de, message = FALSE}


current.de_stat = as.data.frame(de_stat %>% group_by(reduced_dim_method , reduced_dim_type , reduced_dim_genes , reduced_dim_type_genes , reduced_dim , SYMBOL , round) %>% dplyr::summarise(logFC = abs(mean(logFC , na.rm = T)) , logFC_all = mean(logFC_all)))


plots = lapply(levels(current.de_stat$SYMBOL) , function(gene){
  p = ggplot(current.de_stat[current.de_stat$SYMBOL == gene, ] , aes(x = reduced_dim_method , y = logFC , fill = reduced_dim_type_genes)) +
    geom_boxplot(outlier.size = .25 , fatten = 1 , lwd = .2) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    ggtitle(gene) + 
    scale_fill_manual(values = cols , name = "Embedding") + 
    labs(x = "" , y = "") +
    geom_hline(yintercept = abs(unique(current.de_stat$logFC_all[current.de_stat$SYMBOL == gene])) , "color" = "red" , "linetype" = "dashed") +
    theme(legend.position = "none") 
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(chimera_speicific_genes) , nrow = 1)
p
ggsave(filename = paste0(figures.dir, "logFC_distr_combined", ".png"), plot = p, width = 6.5, height = 2.8)




```

### Frac of hoods sig

```{r boxplot-distr-milo-de, message = FALSE}


current.de_stat = as.data.frame(de_stat %>% group_by(SYMBOL , id , reduced_dim_method , reduced_dim_type , reduced_dim_genes , reduced_dim_type_genes , reduced_dim , round) %>% dplyr::summarise(frac_hoods_sig_01 = mean(pval_corrected_across_nhoods < 0.1 , na.rm = T) , frac_hoods_sig_005 = mean(pval_corrected_across_nhoods < 0.05 , na.rm = T) , frac_hoods_sig_001 = mean(pval_corrected_across_nhoods < 0.01 , na.rm = T) , frac_hoods_sig_0001 = mean(pval_corrected_across_nhoods < 0.005 , na.rm = T)  , frac_hoods_sig_0005 = mean(pval_corrected_across_nhoods < 0.001 , na.rm = T) , frac_hoods_sig_00001 = mean(pval_corrected_across_nhoods < 0.0001 , na.rm = T) ))


plots = lapply(levels(chimera_speicific_stat$SYMBOL) , function(gene){
  p = ggplot(current.de_stat[current.de_stat$SYMBOL == gene , ] , aes(x = reduced_dim_method , y = frac_hoods_sig_01 , fill = reduced_dim_type_genes)) + 
    geom_boxplot(outlier.size = .25 , fatten = 1 , lwd = .2) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    labs(x = "" , y = "")  +
    scale_fill_manual(values = cols, name = "Embedding") +
    theme(legend.position = "none") + ggtitle(gene) +
    ylim(c(0.89,1))
  return(p)
})
p = ggarrange(plotlist = plots , common.legend = F , ncol = length(levels(chimera_speicific_stat$SYMBOL)) , nrow = 1)
p
ggsave(filename = paste0(figures.dir, "frac_sig_distr_combined", ".png"), plot = p, width = 6.5, height = 2.8)



```



# Session Info

```{r sessinf}
sessionInfo()
```
