---
title: "Analysis for simulations on brain."
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---
# Load dependencies

```{r load, message = FALSE}


library(SingleCellExperiment)
library(miloDE)
library(miloR)
library(ggplot2)
library(ggpubr)
library(viridis)
library(MetBrewer)
library(BiocParallel)
library(slingshot)

ncores = 5
mcparam = MulticoreParam(workers = ncores)
register(mcparam)

set.seed(32)

root.dir = "/nfs/research/marioni/alsu/hubmap_metaRef/"
figures.dir = paste0(root.dir , "figures/detection_analysis/chimera_embryo/cardiomyocytes/")


# load data
sce = readRDS(paste0(root.dir , "data/sces/mouse_embryo/chimera_WT/mouse_embryo_wt_chimera.Rds"))
sce = sce[, sce$celltype == "Cardiomyocytes"]
set.seed(32)
slingshot_res <- slingPseudotime(slingshot(as.matrix(reducedDim(sce , "mnn_unsupervised_wt"))))
sce$pt = as.numeric(slingshot_res)
pt_grid = sapply(seq(0,1,0.01) , function(i) return(as.numeric(quantile(sce$pt , i))))

cluster_splits = list(c(pt_grid[1] , pt_grid[3]) , 
                      c(pt_grid[49] , pt_grid[51]),
                      c(pt_grid[99] , pt_grid[101]),
                      c(pt_grid[1] , pt_grid[6]) , 
                      c(pt_grid[47] , pt_grid[53]),
                      c(pt_grid[95] , pt_grid[101]),
                      c(pt_grid[1] , pt_grid[11]) , 
                      c(pt_grid[45] , pt_grid[55]),
                      c(pt_grid[91] , pt_grid[101]),
                      c(pt_grid[1] , pt_grid[26]) , 
                      c(pt_grid[37] , pt_grid[63]),
                      c(pt_grid[75] , pt_grid[101])
                      )

simulated_genes = readRDS(file = paste0(root.dir , "data/processed/simulations/cardiomyocytes/simulated_genes.Rds"))
sces = readRDS(paste0(root.dir , "data/processed/simulations/cardiomyocytes/sces_simulated.Rds"))

milo_de_stat = readRDS(paste0(root.dir , "data/processed/simulations/cardiomyocytes/milo_de_stat__recalculated_embedding.Rds"))
bulk_de_stat = readRDS(paste0(root.dir , "data/processed/simulations/cardiomyocytes/bulk_de_stat.Rds"))
bulk_de_stat_w_pt = readRDS(paste0(root.dir , "data/processed/simulations/cardiomyocytes/bulk_de_stat_w_pt.Rds"))


```

# Unlist miloDE and bulkDE with assigning % of cells perturbed

```{r unlist-clusts, message = FALSE}


cluster_splits_df = lapply(1:length(cluster_splits) , function(i){
  out = round( mean(sce$pt >= cluster_splits[[i]][1] & sce$pt <= cluster_splits[[i]][2]) , 3 )
  out = data.frame(idx = i , prop = out)
  return(out)
})
cluster_splits_df = do.call(rbind , cluster_splits_df)



bulk_de_stat = lapply(1:length(bulk_de_stat) , function(i){
  current.stat = bulk_de_stat[[i]]
  current.stat_w_pt = bulk_de_stat_w_pt[[i]]
  current.stat = merge(current.stat , current.stat_w_pt)
  
  current.stat$logFC_bulk_all = -1*(current.stat$logFC_bulk_all)
  current.stat$logFC_bulk_clusters = -1*(current.stat$logFC_bulk_clusters)
  current.stat$logFC_bulk_all_w_pt = -1*(current.stat$logFC_w_pt)
  
  current.stat$logFC_bulk_all_round = round(current.stat$logFC_bulk_all , 1)
  current.stat$logFC_bulk_clusters_round = round(current.stat$logFC_bulk_clusters , 1)
  current.stat$logFC_bulk_all_w_pt_round = round(current.stat$logFC_bulk_all_w_pt)
  
  current.stat = cbind(current.stat , cluster_splits_df[i,])
  return(current.stat)
})
bulk_de_stat = do.call(rbind , bulk_de_stat)


milo_de_stat = lapply(1:length(milo_de_stat) , function(i){
  current.stat = milo_de_stat[[i]]
  current.stat = cbind(current.stat , cluster_splits_df[i,])
  return(current.stat)
})
milo_de_stat = do.call(rbind , milo_de_stat)


```

# Overview of perturbed cells

```{r load-data, message = FALSE}


umaps = as.data.frame( scater::calculateUMAP(t(reducedDim(sce , "mnn_unsupervised_wt"))) )
reducedDim(sce , paste0("umap_" , "mnn_unsupervised_wt")) = umaps

cols = c("#899DA4" , "#C93312")
names(cols) = c(F,T)


get_plot_per_cluster_split = function(idx){
  current.clusters = cluster_splits[[idx]]
  sce$perturbed = sce$pt >= current.clusters[1] & sce$pt <= current.clusters[2]
  
  meta = cbind(as.data.frame(colData(sce)) , as.data.frame(reducedDim(sce , "umap_mnn_unsupervised_wt")))
  meta = meta[order(meta$perturbed) , ]
  p = ggplot(meta , aes(x = V1 , y = V2 , col = perturbed)) + 
    geom_point(size = .5 , alpha = .9) +
    theme_bw() + 
    scale_color_manual(values = cols) + 
    ggtitle(paste0("Prop. = ", cluster_splits_df$prop[cluster_splits_df$idx == idx]))
  return(p)
}

plots = lapply(c(1:length(sces)) , function(i){
  p = get_plot_per_cluster_split(i)
  return(p)
})
p = ggarrange(plotlist = plots , nrow = 3 , ncol = 4, common.legend = T)
p
ggsave(filename = paste0(figures.dir, "overview_perturbations", ".png"), plot = p, width = 8, height = 5)


```


# bulkDE stat

## Heatmap: lfc x prop

```{r load-data, message = FALSE}


bulk_de_stat_combined = as.data.frame(bulk_de_stat %>% group_by(idx , prop , logFC_bulk_clusters_round) %>% 
                                        dplyr::summarise(frac_genes_sig = mean(PValue_bulk_all < 0.05) , 
                                                         avg_logFC_all = mean(logFC_bulk_all_round)))


p = ggplot(bulk_de_stat_combined[bulk_de_stat_combined$logFC_bulk_clusters < 6 & bulk_de_stat_combined$logFC_bulk_clusters > 0, ] , aes(x = factor(idx) , y = factor(logFC_bulk_clusters_round) , fill = frac_genes_sig)) + 
  geom_tile() + 
  scale_fill_viridis(discrete = F , name = "% genes\ndet. as DE") +
  theme_bw() + 
  labs(x = "Proportion of perturbed cells" , y = "logFC for perturbed cells")
p
ggsave(filename = paste0(figures.dir, "bulk_de/heatmap__det_rate", ".png"), plot = p, width = 5, height = 5)


p = ggplot(bulk_de_stat_combined[bulk_de_stat_combined$logFC_bulk_clusters < 4 & bulk_de_stat_combined$logFC_bulk_clusters > 0, ] , aes(x = factor(prop) , y = factor(logFC_bulk_clusters_round) , fill = avg_logFC_all)) + 
  geom_tile() + 
  scale_fill_viridis(discrete = F, name = "abs\nlogFC") +
  theme_bw() + 
  labs(x = "Proportion of perturbed cells" , y = "logFC for perturbed cells")
p
ggsave(filename = paste0(figures.dir, "bulk_de/heatmap__abs_logFC", ".png"), plot = p, width = 5, height = 5)


```

## compare logFC (abs vs perturbed)

```{r load-data, message = FALSE}


cols = wes_palette("Zissou1", length(unique(bulk_de_stat$prop)), type = "continuous")

df = bulk_de_stat[bulk_de_stat$logFC_bulk_clusters < 4 & bulk_de_stat$logFC_bulk_clusters > 0 ,]
df = df[sample(nrow(df)) , ]
p = ggplot(df, aes(y = logFC_bulk_all , x = logFC_bulk_clusters , col = factor(prop))) +
  geom_point(alpha = .75 , size = .75) +
  theme_bw() +
  scale_color_manual(values = cols[1:length(unique(bulk_de_stat$prop))] , name = "Prop. of\nperturbed\ncells") + 
  geom_abline(slope = 1 , "linetype" = "dashed" , color = "black") + 
  labs(x = "Estimated logFC in perturbed sub-population" , y = "logFC from DE test") 
p
ggsave(filename = paste0(figures.dir, "bulk_de/scater_logFC", ".png"), plot = p, width = 5, height = 4)





```

# miloDE - sensitivity of DE detection

## Get stat

For each simulation, nhood assignment, gene and CT purity thresh - get sensitivity and FDR

```{r load-data, message = FALSE}


ct_purity.thresh = c(0, 0.1 , 0.25 , 0.5)

milo_de_stat_combined = lapply(ct_purity.thresh , function(ct_purity){
  out = as.data.frame(milo_de_stat %>% group_by(idx , prop , order , k , round , id , ENSEMBL , SYMBOL ) %>% 
                        dplyr::summarise(sensitivity = sum(pval < 0.05 & cluster_purity > ct_purity)/sum(cluster_purity > ct_purity) , 
                                         fdr = sum(pval < 0.05 & cluster_purity <= ct_purity)/sum(pval < 0.05) ))
  out$ct_purity.thresh = ct_purity
  return(out)
})
milo_de_stat_combined = do.call(rbind, milo_de_stat_combined)
milo_de_stat_combined = merge(milo_de_stat_combined , bulk_de_stat[, c("ENSEMBL" , "idx" , "logFC_bulk_clusters_round")] , by = c("ENSEMBL" , "idx") , all.x = T , all.y = F)


milo_de_stat_combined = as.data.frame(milo_de_stat_combined %>% group_by(idx , prop ,logFC_bulk_clusters_round , ct_purity.thresh) %>% 
                                        dplyr::summarise(sensitivity = mean(sensitivity , na.rm = T) , fdr = mean(fdr , na.rm = T)))


plots = lapply(unique(milo_de_stat_combined$ct_purity.thresh) , function(ct_purity){
  current.stat = milo_de_stat_combined[milo_de_stat_combined$logFC_bulk_clusters_round < 6 & milo_de_stat_combined$logFC_bulk_clusters_round > 0 & milo_de_stat_combined$ct_purity.thresh == ct_purity, ]
  
  p = ggplot(current.stat , aes(x = factor(prop) , y = factor(logFC_bulk_clusters_round) , fill = sensitivity)) + 
    geom_tile() + 
    scale_fill_viridis(discrete = F , name = "Sensitivity") +
    theme_bw() +
    labs(x = "Proportion of perturbed cells" , y = "logFC for perturbed cells")
  p
  ggsave(filename = paste0(figures.dir, "milo_de/heatmaps_sensitivity/ct_purity_", ct_purity , ".png"), plot = p, width = 5, height = 5)
  
  p = ggplot(current.stat , aes(x = factor(prop) , y = factor(logFC_bulk_clusters_round) , fill = fdr)) + 
    geom_tile() + 
    scale_fill_viridis(discrete = F , name = "FDR") +
    theme_bw() +
    labs(x = "Proportion of perturbed cells" , y = "logFC for perturbed cells")
  p
  ggsave(filename = paste0(figures.dir, "milo_de/heatmaps_fdr/ct_purity_", ct_purity , ".png"), plot = p, width = 5, height = 5)
})



```

## Compare logFC (abs vs perturbed)

```{r load-data, message = FALSE}


cols = wes_palette("Zissou1", length(unique(bulk_de_stat$prop)), type = "continuous")


df = merge(milo_de_stat , bulk_de_stat[, c("ENSEMBL" , "idx" , "logFC_bulk_clusters")] , by = c("ENSEMBL" , "idx") , all.x = T , all.y = F)
df = df[df$logFC_bulk_clusters < 4 & df$logFC_bulk_clusters > 0 ,]
df$logFC = -1*df$logFC 
df = df[sample(nrow(df)) , ]


p1 = ggplot(df[df$cluster_purity > 0.1 & df$pval_corrected_across_nhoods < 0.1, ], aes(y = logFC , x = logFC_bulk_clusters , col = factor(prop))) +
  geom_point(alpha = .75 , size = .75) +
  theme_bw() +
  scale_color_manual(values = cols[1:length(unique(bulk_de_stat$prop))] , name = "Prop. of\nperturbed\ncells") + 
  geom_abline(slope = 1 , "linetype" = "dashed" , color = "black") + 
  labs(x = "Estimated logFC in perturbed sub-population" , y = "logFC from DE test") 
p2 = ggplot(df[df$cluster_purity > 0.1 & df$pval_corrected_across_nhoods < 0.1, ], aes(y = logFC , x = logFC_bulk_clusters , col = cluster_purity)) +
  geom_point(alpha = .75 , size = .75) +
  theme_bw() +
  geom_abline(slope = 1 , "linetype" = "dashed" , color = "black") + 
  scale_color_viridis(discrete = F , option = "cividis" , name = "Nhood\npurity") +
  labs(x = "Estimated logFC in perturbed sub-population" , y = "logFC from DE test") 
p = ggarrange(p1,p2)
p
ggsave(filename = paste0(figures.dir, "milo_de/scater_logFC", ".png"), plot = p, width = 10, height = 4)





```


## Robustness - based on k and on replicate of the simulation

@@ talk to John what is the best way to show it

```{r load-data, message = FALSE}






```

# Session Info

```{r sessinf}
sessionInfo()
```
