---
title: "This script is to assess how DE detection changes with number of cells / sample composition."
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---
# Load dependencies

```{r load, message = FALSE}


library(SingleCellExperiment)
library(batchelor)
library(stats)
library(Matrix)
library(splatter)
library(BiocParallel)
library(BiocNeighbors)
library(BiocSingular)
library(pROC)
library(ggplot2)
library(ggpubr)
library(scran)
library(wesanderson)
library(stringr)
library(MetBrewer)

set.seed(32)

#root.dir = "/Users/alsu/Develop/hubmap_metaRef/"
root.dir = "/nfs/research/marioni/alsu/hubmap_metaRef/"
figures.dir = paste0(root.dir , "figures/DE_detection_n_cells/")
#source(paste0(root.dir , "am_hubmapMetaRef/functions/milo_functions.R"))
#source(paste0(root.dir , "am_hubmapMetaRef/functions/core_functions.R"))     

vars = c("sensitivity" , "specificity" , "fdr")


```

# Add stat on samples composition (n-total and ratio)

```{r sub-ds-analysis, message = FALSE}


add_sample_stat_to_de = function(de_stat){
  cols_group1 = paste0( paste0("Batch" , c(1:10)) , "_Group1")
  cols_group2 = paste0( paste0("Batch" , c(1:10)) , "_Group2")
  de_stat$n_samples_total_actual = sapply(1:nrow(de_stat) , function(i){
    return(sum(de_stat[i,c(cols_group1 , cols_group2)] != 0))
  })
  de_stat$n_samples_ratio_actual = sapply(1:nrow(de_stat) , function(i){
    out = sum(de_stat[i,cols_group1] != 0)/sum(de_stat[i,cols_group2] != 0)
    return( round(out,2) )
  })
  de_stat$n_samples_total = de_stat$n_samples_ref + de_stat$n_samples_query
  de_stat$n_samples_ratio = sapply(1:nrow(de_stat) , function(i){
    out = de_stat$n_samples_ref[i]/de_stat$n_samples_query[i]
    return( round(out,2) )
  })
  de_stat$avg_n_cells_downsampled_per_sample = de_stat$n_cells_downsampled / de_stat$n_samples_total
  return(de_stat)
}

# original
de_stat_downsampling = readRDS(paste0(root.dir , "data/processed/DE_detection_n_cells/de_stat_downsampling.Rds"))
de_stat_downsampling = add_sample_stat_to_de(de_stat_downsampling)
# add ds ID
de_stat_downsampling$dataset = paste0(de_stat_downsampling$lfc , "_" , de_stat_downsampling$n_samples_ref , "_" , de_stat_downsampling$n_samples_query)
de_stat_downsampling$id = paste0(de_stat_downsampling$dataset , "_lfc.thresh_" , de_stat_downsampling$lfc.thresh)

# batch
de_stat_downsampling_batch = readRDS(paste0(root.dir , "data/processed/DE_detection_n_cells_batch/de_stat_downsampling.Rds"))
de_stat_downsampling_batch = add_sample_stat_to_de(de_stat_downsampling_batch)
# add ds ID
de_stat_downsampling_batch$dataset = paste0(de_stat_downsampling_batch$lfc , "_" , de_stat_downsampling_batch$n_samples_ref , "_" , de_stat_downsampling_batch$n_samples_query)
de_stat_downsampling_batch$id = paste0(de_stat_downsampling_batch$dataset , "_lfc.thresh_" , de_stat_downsampling_batch$lfc.thresh)



```

# Add runmed for each dataset (logFC + n-samples-ref + n-samples-query)

```{r run-med, message = FALSE}


## add runmed
add_runmed = function(de_stat_downsampling , var_y , var_x = "n_cells_downsampled"){
  require(stats)
  de_stat_downsampling = de_stat_downsampling[order(de_stat_downsampling[ , var_x]) , ]
  de_stat_downsampling[, paste0(var_y , "_runmed_" , var_x)] = NaN
  for (id in unique(de_stat_downsampling$id) ){
    idx = which(de_stat_downsampling$id == id )
    current.runmed = runmed(de_stat_downsampling[idx, var_y] , k = 51 , endrule = "constant")
    de_stat_downsampling[idx,paste0(var_y , "_runmed_" , var_x)] = current.runmed
  }
  return(de_stat_downsampling)
}

for (var in vars){
  de_stat_downsampling = add_runmed(de_stat_downsampling , paste0(var) , var_x = "n_cells_downsampled")
  de_stat_downsampling = add_runmed(de_stat_downsampling , paste0(var) , var_x = "avg_n_cells_downsampled_per_sample")
  
  de_stat_downsampling_batch = add_runmed(de_stat_downsampling_batch , paste0(var) , var_x = "n_cells_downsampled")
  de_stat_downsampling_batch = add_runmed(de_stat_downsampling_batch , paste0(var) , var_x = "avg_n_cells_downsampled_per_sample")
}


```

# How do sensitivity, specificity and FDR change 

Fix lfc = 1 datasets

## Set plotting functions

```{r get-plot-func, message = FALSE}


get_runmed_plot_one_var_col = function(df , var , cols , col_factor , var_x , size = 0.3){
  df$col_factor = factor( df[,col_factor] , levels = unique(sort(df[,col_factor])))
  p = ggplot(df, aes(x = df[, var_x] , y = df[, var] , col = col_factor, group = factor(id))) + 
    geom_line(size=size) + 
    #labs(y = str_remove( var , "_runmed_")  , x = var_x) + 
    theme_bw() + 
    scale_color_manual(values = cols , name = "") + 
    labs(x = "" , y = "") +
    theme(axis.text=element_text(size=13)) +
    theme(legend.text=element_text(size=11) , legend.title = element_text(size=11)) + 
    theme(text=element_text(size=11) , axis.text.x = element_text(size=11) , axis.text.y = element_text(size=11)) +
    guides(linetype = guide_legend(override.aes = list(size = 2)))
  return(p)
}



```


## Equal number of samples - different n-samples

```{r dynamic-n-cells-n-samples-equal, message = FALSE}


datasets = c("1_2_2" , "1_4_4" , "1_6_6" , "1_8_8" , "1_10_10")
df_lfc.thresh_0 = de_stat_downsampling[de_stat_downsampling$dataset %in% datasets & de_stat_downsampling$lfc.thresh == 0 , ]
df_lfc.thresh_1 = de_stat_downsampling[de_stat_downsampling$dataset %in% datasets & de_stat_downsampling$lfc.thresh == 0.5 , ]
col_factor = "n_samples_total"
cols = wes_palette("Zissou1" , n = 5)
var_x_n_cells_downsampled = "n_cells_downsampled"
var_x_avg_n_cells_downsampled_per_sample = "avg_n_cells_downsampled_per_sample"


p1_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.4 , 0.9)) +
  geom_rect(aes(xmin = 200, xmax = 700, ymin = 0.55, ymax = 0.77) , alpha = 0 , "linetype" = "dashed" , col = "gray20" , size = 0.3)
p2_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "specificity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.92 , 0.99))
p3_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "fdr_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.05 , 0.35))

p1_2 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.4 , 0.9))
p2_2 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "specificity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.92 , 0.99))
p3_2 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "fdr_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.05 , 0.35))

p1_3 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.4 , 0.9))
p2_3 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "specificity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.92 , 0.99))
p3_3 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "fdr_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.05 , 0.35))

p1_4 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "sensitivity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.4 , 0.9))
p2_4 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "specificity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.92 , 0.99))
p3_4 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "fdr_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.05 , 0.35))


p = ggarrange(p1_1, p1_2, p1_3, p1_4, p2_1, p2_2, p2_3, p2_4 , p3_1, p3_2, p3_3, p3_4 , nrow = 3 , ncol = 4, common.legend = T) + theme(legend.position = "right")
p
ggsave(filename = paste0(figures.dir , "stat_equal_n_samples.png"), plot = p, width = 12, height = 7)



# get legend
p = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled, size = 1.5) +
  theme(legend.position = "right")
p
ggsave(filename = paste0(figures.dir , "stat_equal_n_samples_legend.png"), plot = p, width = 5, height = 5)


# inlet to zoom in
p1_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(200,700)) + ylim(c(0.55 , 0.77)) +
  theme(legend.position = "none") +
  labs(x = "Total number of cells" , y = "Sensitivity") 
ggsave(filename = paste0(figures.dir , "stat_equal_n_samples_inlet.png"), plot = p1_1, width = 3, height = 3)



```

## Equal n-samples - different n-samples

```{r dynamic-n-cells-n-samples-diff, message = FALSE}


datasets = c("1_2_10" , "1_3_9" , "1_4_8" , "1_6_6" , "1_8_4" , "1_9_3" , "1_10_2")
df_lfc.thresh_0 = de_stat_downsampling[de_stat_downsampling$dataset %in% datasets & de_stat_downsampling$lfc.thresh == 0 , ]
df_lfc.thresh_1 = de_stat_downsampling[de_stat_downsampling$dataset %in% datasets & de_stat_downsampling$lfc.thresh == 0.5 , ]
col_factor = "n_samples_ratio"
cols = c("aquamarine4" , "deeppink2" , "goldenrod2" , "black" , "khaki1" , "lightpink2" , "aquamarine2")
var_x_n_cells_downsampled = "n_cells_downsampled"
var_x_avg_n_cells_downsampled_per_sample = "avg_n_cells_downsampled_per_sample"



p1_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.4 , 0.9))
p2_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "specificity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.92 , 0.99))
p3_1 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "fdr_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.05 , 0.35))

p1_2 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.4 , 0.9))
p2_2 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "specificity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.92 , 0.99))
p3_2 = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "fdr_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.05 , 0.35))

p1_3 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.4 , 0.9))
p2_3 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "specificity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.92 , 0.99))
p3_3 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "fdr_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled) + xlim(c(100,2500)) + ylim(c(0.05 , 0.35))

p1_4 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "sensitivity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.4 , 0.9))
p2_4 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "specificity_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.92 , 0.99))
p3_4 = get_runmed_plot_one_var_col(df_lfc.thresh_1 , var = "fdr_runmed_avg_n_cells_downsampled_per_sample" , cols , col_factor = col_factor , var_x = var_x_avg_n_cells_downsampled_per_sample) + xlim(c(5,150)) + ylim(c(0.05 , 0.35))


p = ggarrange(p1_1, p1_2, p1_3, p1_4, p2_1, p2_2, p2_3, p2_4 , p3_1, p3_2, p3_3, p3_4 , nrow = 3 , ncol = 4, common.legend = T) + theme(legend.position = "right")
p
ggsave(filename = paste0(figures.dir , "stat_diff_ratio.png"), plot = p, width = 12, height = 7)


# get legend
p = get_runmed_plot_one_var_col(df_lfc.thresh_0 , var = "sensitivity_runmed_n_cells_downsampled" , cols , col_factor = col_factor , var_x = var_x_n_cells_downsampled, size = 1.5) +
  theme(legend.position = "right")
p
ggsave(filename = paste0(figures.dir , "stat_diff_ratio_legend.png"), plot = p, width = 5, height = 5)



```

## Different logFC & batch effect - just sensitivity

```{r get-sens-diff-lfc, message = FALSE}



cols = met.brewer("Hokusai2" , n = 8)[c(1,3,6,8)]
df_1 = de_stat_downsampling[de_stat_downsampling$lfc.thresh == 0 & de_stat_downsampling$lfc > 0.5 & de_stat_downsampling$n_cells_downsampled < 1000 & de_stat_downsampling$n_cells_downsampled > 100, ]
df_1$batch = "Original (no batch effect)"
df_2 = de_stat_downsampling_batch[de_stat_downsampling_batch$lfc.thresh == 0 & de_stat_downsampling_batch$lfc > 0.5 & de_stat_downsampling_batch$n_cells_downsampled < 1000 & de_stat_downsampling_batch$n_cells_downsampled > 100, ]
df_2$batch = "With batch effect"
df = rbind(df_1 , df_2)
df$batch_lfc = paste0(df$batch , ",\nlogFC = ", df$lfc)
cols = wes_palette("Zissou1" , n = 5)

p = ggplot(df[df$n_samples_ratio == 1 & df$lfc <= 3, ], aes(x = n_cells_downsampled , y = sensitivity_runmed_n_cells_downsampled , col = factor(n_samples_ref) , group = factor(id))) + 
  geom_line(size=0.5) + 
  theme_bw() + 
  facet_wrap(~batch_lfc) +
  labs(x = "" , y = "") +
  theme(axis.text=element_text(size=13)) +
  theme(legend.text=element_text(size=11) , legend.title = element_text(size=11) ) + 
  theme(text=element_text(size=11) , axis.text.x = element_text(size=11) , axis.text.y = element_text(size=11)) +
  scale_color_manual(values = cols , name = "") + theme(legend.position = "none")
  #scale_x_continuous(limits = c(0, 1500), breaks = c(0 , 500 , 1000))
p
ggsave(filename = paste0(figures.dir , "stat_diff_lfc_w_batch.png"), plot = p, width = 8, height = 5)


```

# Save ds

```{r save-ds, message = FALSE}

df = de_stat_downsampling[de_stat_downsampling$n_samples_total == de_stat_downsampling$n_samples_total_actual & de_stat_downsampling$lfc > 0.5, c("n_samples_ref" , "n_samples_query" , "lfc" , "n_cells_downsampled" , "avg_n_cells_downsampled_per_sample", "n_samples_total" , "n_samples_ratio" , "lfc.thresh" , "sensitivity_runmed_n_cells_downsampled" , "specificity_runmed_n_cells_downsampled" , "fdr_runmed_n_cells_downsampled" , "id")]
df = df[df$n_cells_downsampled < 2000 , ]

# get grid 
n_cells.grid = seq(50 , 2000 , 5)

get_selection_per_grid = function(id , n_cells.grid){
  current.df = df[df$id == id , ]
  current.df = current.df[order(current.df$n_cells_downsampled) , ]
  idx_keep = sapply(n_cells.grid , function(n_cells){
    out = abs(current.df$n_cells_downsampled - n_cells)
    idx = which(out == min(out))
    idx = idx[1]
    return(idx)
  })
  idx_keep = unique(idx_keep)
  current.df = current.df[idx_keep , ]
  return(current.df)
}


ids = unique(df$id)
stat = lapply(ids , function(id){
  out = get_selection_per_grid(id , n_cells.grid)
  return(out)
})
stat = do.call(rbind , stat)
stat = stat[ , c("lfc" , "n_samples_ref" , "n_samples_query" , "n_samples_total" , "n_samples_ratio" , "n_cells_downsampled" , "avg_n_cells_downsampled_per_sample","lfc.thresh","sensitivity_runmed_n_cells_downsampled" , "specificity_runmed_n_cells_downsampled" , "fdr_runmed_n_cells_downsampled")]
write.csv(stat , file = paste0(figures.dir , "stat.csv") , row.names = F , col.names = T)


```


# Session Info

```{r sessinf}
sessionInfo()
```
