---
title: "Analysis of simulated data for order analysis."
output:
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---
# Load dependencies

```{r load, message = FALSE}


library(miloDE)
library(SingleCellExperiment)
library(batchelor)
library(tibble)
library(batchelor)
library(stringr)
library(igraph)
library(miloR)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(BiocParallel)
library(ggbeeswarm)
library(cluster)

ncores = 5
mcparam = MulticoreParam(workers = ncores)
register(mcparam)

set.seed(32)

#root.dir = "/Users/alsu/Develop/hubmap_metaRef/"
root.dir = "/nfs/research/marioni/alsu/hubmap_metaRef/"
figures.dir = paste0(root.dir , "figures/sensitivity_analysis_full_ds/chimera_embryo/")

cols_order = c("#3b7c70", "#ce9642")
names(cols_order) = c(1,2)


```


# Load simulated genes, sce-milo and milo-de

```{r load-data, message = FALSE}


simulated_genes = readRDS(file = paste0(root.dir , "data/processed/order_analysis_full_ds/simulated_genes_full_ds.Rds"))

anno_1 = expand.grid(order = 1, k = c(75,150,200,250,300,350,400,450) , prop = c(0.2) , round = c(1:10))
anno_2 = expand.grid(order = 2, k = seq(10,45,5), prop = c(0.2), round = c(1:10))
anno = rbind(anno_1 , anno_2)

# sort out sub-celltypes
sce = readRDS(paste0(root.dir , "data/processed/order_analysis_full_ds/sce_chimera_w_simulations_full_ds.Rds"))
# update type names
#sce$type = sapply(sce$type , function(x) ifelse(x == "wt" , "WT" , "ChimeraWT"))
#sce$type = factor(sce$type , levels = c("WT" , "ChimeraWT"))

rowdata = as.data.frame(rowData(sce))

ct_levels_full = names(sort(table(sce$celltype)))
ct_levels = ct_levels_full[ct_levels_full %in% unique(simulated_genes$celltype)]


# get original CT naming
sce_original = readRDS(paste0(root.dir , "data/sces/mouse_embryo/chimera/mouse_embryo_chimera_tal.Rds"))
ct_anno = data.frame(celltype_proper = unique(sce_original$celltype))
celltypes = sapply(ct_anno$celltype_proper , function(celltype) str_replace_all(celltype , " " , "_"))
celltypes = sapply(celltypes , function(celltype) str_replace_all(celltype , "/" , "_"))
ct_anno$celltype = celltypes
ct_anno = ct_anno[ct_anno$celltype %in% ct_levels_full , ]

celltypes_proper = sapply(sce$celltype , function(celltype) return(ct_anno$celltype_proper[ct_anno$celltype == celltype]))
#celltypes_proper = unlist(celltypes_proper)
sce$celltype_proper = celltypes_proper
#saveRDS(sce , paste0(root.dir , "data/processed/order_analysis_full_ds/sce_chimera_w_simulations_full_ds.Rds"))


```


# Get UMAP

```{r get-umap, message = FALSE}


set.seed(20)
umaps = as.data.frame(uwot::umap(reducedDim(sce , "scVI") , min_dist = 0.7))
umaps = cbind(umaps , as.data.frame(colData(sce)))
umaps = umaps[umaps$celltype %in% ct_levels , ]
umaps$celltype_proper = as.character(umaps$celltype_proper)
umaps$type = factor(umaps$type , levels = c("WT" , "ChimeraWT"))
p = ggplot(umaps , aes(x = V1 , y = V2 , col = celltype_proper)) +
  geom_point(size = 0.15) +
  scale_color_manual(values = EmbryoCelltypeColours , name = "Cell type") +
  theme_bw() + guides(colour = guide_legend(override.aes = list(size=4))) +
  labs(x = "UMAP-1" , y = "UMAP-2") + 
  facet_wrap(~type)
p  
ggsave(filename = paste0(figures.dir, "umaps", ".png"), plot = p, width = 11, height = 3.75)


```


# CT composition

```{r ct-composition, message = FALSE}



tab = as.data.frame(table(sce$celltype , sce$type))
colnames(tab) = c("celltype" , "type" , "n")
tab$frac = tab$n/sum(tab$n)
tab$n = log10(tab$n )
tab = tab[tab$celltype %in% ct_levels , ]
#tab$celltype = factor(tab$celltype , levels = ct_levels)
tab = merge(tab , ct_anno , all.x = T , all.y = F)
# order
tab_order = as.data.frame(table(sce$celltype_proper[sce$celltype_proper %in% ct_anno$celltype_proper]))
colnames(tab_order) = c("celltype_proper" , "n")
tab_order = tab_order[order(tab_order$n) , ]
tab$celltype_proper = factor(tab$celltype_proper , levels = tab_order$celltype_proper)

tab = melt(tab[,c("celltype_proper" , "type" , "n" , "frac")] , id = c("celltype_proper" , "type"))
tab$celltype_proper = factor(tab$celltype_proper , levels = tab_order$celltype_proper)

cols = wesanderson::wes_palette("Royal1" , n = 2)
tab$variable = sapply(tab$variable , function(x) ifelse(x == "n" , "log10(#cells)" , "Fraction of cells"))
tab$type = sapply(tab$type , function(x) ifelse(x == "wt" , "WT" , "ChimeraWT"))
tab$type =  factor(tab$type , levels = c("WT" , "ChimeraWT"))


p = ggplot(tab , aes(x = celltype_proper , y = value, fill = type)) +
  geom_bar(stat = "identity" , position = "dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  scale_fill_manual(values = cols , name = "Condition") +
  labs(x = "Cell type" , y = "") +
  facet_wrap(~variable , scales = "free_y" , nrow = 2) +
  #ylim(c(0, 200)) + 
  theme(axis.text=element_text(size=13)) +
  theme(legend.text=element_text(size=13) , legend.title = element_text(size=13) ) + 
  theme(text=element_text(size=13) , axis.text.x = element_text(size=13) , axis.text.y = element_text(size=13)) 
p
ggsave(filename = paste0(figures.dir, "ct_composition", ".png"), plot = p, width = 9, height = 6)




```


## What is logFC estimate for our genes

```{r estimate-logfc, message = FALSE}


get_de_genes = function( celltype , genes){
  require(edgeR)
  current.sce = sce[, sce$celltype == celltype]
  summed = summarizeAssayByGroup(counts(current.sce), colData(current.sce)[,c("sample","type")])
  y <- DGEList(assay(summed, "sum"), samples=colData(summed), lib.size = colSums(assay(summed , "sum")))
  y <- calcNormFactors(y)
  design <- model.matrix(~ factor(type), y$samples)
  y <- estimateDisp(y, design)
  fit <- glmQLFit(y, design, robust=TRUE)
  res <- glmQLFTest(fit, coef=ncol(design))
  out = topTags(res, n = Inf )
  out = as.data.frame(out)
  out = rownames_to_column(out , var = "ENSEMBL")
  out = merge(out , as.data.frame(rowData(sce)))
  out$celltype = celltype
  out = out[out$ENSEMBL %in% genes , ]
  return(out)
}


de_stat = lapply(unique(simulated_genes$celltype) , function(celltype){
  current.genes = simulated_genes$ENSEMBL[simulated_genes$celltype == celltype]
  out = get_de_genes(celltype , current.genes)
  return(out)
})
de_stat = do.call(rbind , de_stat)
simulated_genes = merge(de_stat[ , c("ENSEMBL" , "SYMBOL", "celltype" , "logFC")] , simulated_genes , by = c("ENSEMBL" , "SYMBOL", "celltype"))
simulated_genes$logFC_round = round(simulated_genes$logFC , 1 )
colnames(simulated_genes)[colnames(simulated_genes) == "logFC"] = "logFC_bulk"


```


# Get homogeneity stat

## Get CT purity stat

```{r ct-purity-stat, message = FALSE}


get_ct_stat_single_milo = function(sce_milo , celltype){
  nhoods_sce = nhoods(sce_milo)
  # celltype specificity per hood
  cells_ct = which(sce_milo$celltype == celltype)
  stat = lapply(1:ncol(nhoods_sce), function(i){
    cells_hood = as.numeric(which(nhoods_sce[,i] == 1))
    out = data.frame(ct_specificity = length(intersect(cells_hood , cells_ct))/length(cells_hood), ct_n_cells = length(intersect(cells_hood , cells_ct)) , 
                     ct_inclusion = length(intersect(cells_hood , cells_ct))/length(cells_ct))
    return(out)
  })
  stat = do.call(rbind , stat)
  stat$Nhood = 1:ncol(nhoods_sce)
  stat$hood_size = colSums(nhoods_sce)
  stat$celltype = celltype
  
  # add average hood size
  stat$avg_hood_size = round( mean(colSums(nhoods_sce)) )
  return(stat)
}



stat_ct_specificity = bplapply(1:nrow(anno) , function(i){
  print(i)
  sce_milo = readRDS(paste0(root.dir ,  "data/processed/order_analysis_full_ds/sce_milo/" ,"sce_order_", anno$order[i] , "_k_" , anno$k[i] , "_round_" , anno$round[i], ".Rds"))
  stat_per_ct = lapply(ct_levels , function(celltype){
    out = get_ct_stat_single_milo(sce_milo , celltype)
    return(out)
  })
  stat_per_ct = do.call(rbind , stat_per_ct)
  stat_per_ct = cbind(stat_per_ct , anno[i, ])
  stat_per_ct$idx = i
  return(stat_per_ct)
}, BPPARAM = mcparam)
stat_ct_specificity = do.call(rbind , stat_ct_specificity)
stat_ct_specificity = stat_ct_specificity[order(stat_ct_specificity$avg_hood_size) , ]
stat_ct_specificity$celltype = factor(stat_ct_specificity$celltype , levels = ct_levels)
stat_ct_specificity$order = factor(stat_ct_specificity$order , levels = c(1,2))
# add order-k as id for coverage
stat_ct_specificity$order_k = paste0(stat_ct_specificity$order , "_" , stat_ct_specificity$k)
saveRDS(stat_ct_specificity , paste0(root.dir , "data/processed/order_analysis_full_ds/nhood_stat_ct_specificity.Rds"))
stat_ct_specificity = readRDS(paste0(root.dir , "data/processed/order_analysis_full_ds/nhood_stat_ct_specificity.Rds"))


```

## Get cell tpye enrichment stat

```{r cell-type-enrichment-stat, message = FALSE}


stat = lapply(1:nrow(anno) , function(i){
  print(i)
  sce_milo = readRDS(paste0(root.dir ,  "data/processed/order_analysis_full_ds/sce_milo/sce_" , "order_" , anno$order[i] , "_k_" , anno$k[i], "_round_", anno$round[i] , ".Rds"))
  nhoods_stat = data.frame(Nhood = 1:ncol(nhoods(sce_milo)))
  nhoods_stat = annotateNhoods(sce_milo , nhoods_stat , coldata_col = "celltype")
  nhoods_stat = merge(nhoods_stat , ct_anno , all.x = T , all.y = F)
  nhoods_stat = cbind(nhoods_stat , anno[i,])
  return(nhoods_stat)
})
stat = do.call(rbind , stat)
stat$id = paste0(stat$order , "_" , stat$k , "_" , stat$round)
stat$order_k = paste0(stat$order , "_" , stat$k)
# only select for tested celltypes
stat = stat[stat$celltype %in% ct_levels , ]
stat$celltype = factor(stat$celltype , levels = ct_levels)
stat$order = factor(stat$order , levels = c(1,2))
saveRDS(stat , paste0(root.dir ,  "data/processed/order_analysis_full_ds/nhood_stat_for_celltype_enrichment.Rds"))
stat_inclusion = readRDS(paste0(root.dir ,  "data/processed/order_analysis_full_ds/nhood_stat_for_celltype_enrichment.Rds"))


```

### Aggregated stat for each coverage - number of nhoods and max CT-specificity

```{r ct-purity-per-assignment, message = FALSE}


tab_specificity = as.data.frame(stat_ct_specificity %>% group_by(celltype , avg_hood_size , order , k , order_k,  prop , round) %>% dplyr::summarise(n_hoods = sum(ct_specificity > 0) , max_ct_specificity = max(ct_specificity) , max_ct_inclusion = max(ct_inclusion)))
tab_inclusion = as.data.frame(stat_inclusion %>% group_by(celltype , order , k , order_k,  prop , round) %>% dplyr::summarise(avg_celltype_fraction = mean(celltype_fraction)))
tab = merge(tab_specificity , tab_inclusion , by = c("celltype" , "order" , "k" , "order_k" , "prop" , "round") , all.x = T)

                              
# add avg-hood-size bin
stat_ct_specificity_per_coverage = as.data.frame(stat_ct_specificity %>% group_by(order_k) %>% dplyr::summarise(avg_hood_size_bin = round( mean(avg_hood_size) )) )
stat_ct_specificity_per_coverage = stat_ct_specificity_per_coverage[order(stat_ct_specificity_per_coverage$avg_hood_size) , ]
tab$order_k = factor(tab$order_k , levels = stat_ct_specificity_per_coverage$order_k)
stat_ct_specificity_per_coverage$order_k = factor(stat_ct_specificity_per_coverage$order_k , levels = stat_ct_specificity_per_coverage$order_k)

stat_ct_specificity_per_coverage$avg_hood_size_bin = factor(stat_ct_specificity_per_coverage$avg_hood_size_bin)
tab = merge(tab , stat_ct_specificity_per_coverage , all.x = T)
# merge w proper cell type names
tab = merge(tab , ct_anno , all.x = T , all.y = F)
tab_order = as.data.frame(table(sce$celltype_proper[sce$celltype_proper %in% ct_anno$celltype_proper]))
colnames(tab_order) = c("celltype_proper" , "n")
tab_order = tab_order[order(tab_order$n) , ]
tab$celltype_proper = factor(tab$celltype_proper , levels = tab_order$celltype_proper[tab_order$celltype_proper %in% tab$celltype_proper])



# plot
## order-k -> nhood size
p = ggplot(tab , aes(x = order_k , y = avg_hood_size , col = order )) +
  geom_boxplot() + 
  scale_color_manual(values = cols_order , name = "Order") +
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "top") + 
  labs(x = "Order-k" , y = "Average\nneighbourhood size" ) +
  theme(axis.text=element_text(size=12)) +
  theme(legend.text=element_text(size=12) , legend.title = element_text(size=12) ) + 
  theme(text=element_text(size=12) , axis.text.x = element_text(size=10) , axis.text.y = element_text(size=12)) 
p
ggsave(filename = paste0(figures.dir, "order_k__hood_size", ".png"), plot = p, width = 5, height = 4)


# avg hood size - number of hoods
p = ggplot(tab , aes(x = avg_hood_size_bin , y = n_hoods , col = order )) +
  geom_boxplot() + geom_jitter(width = 0.2 , size = 0.2) +
  scale_color_manual(values = cols_order , name = "order") +
  theme_bw()  + 
  facet_wrap(~celltype_proper, scales = "free" , ncol = 6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1 )) +
  theme(legend.position = "top") + 
  labs(x = "Average neighbourhood size" , y = "# hoods" ) +
  theme(axis.text=element_text(size=12)) +
  theme(legend.text=element_text(size=12) , legend.title = element_text(size=12) ) + 
  theme(text=element_text(size=12) , axis.text.x = element_text(size=10) , axis.text.y = element_text(size=12))
p
ggsave(filename = paste0(figures.dir, "avg_hood_size__n_hoods", ".png"), plot = p, width = 13, height = 7)


# purity
p = ggplot(tab , aes(x = avg_hood_size_bin , y = max_ct_specificity , col = order )) +
  geom_boxplot() + geom_jitter(width = 0.1 , size = 0.2) +
  scale_color_manual(values = cols_order , name = "Order") +
  theme_bw()  + 
  facet_wrap(~celltype_proper, scales = "free_x",  ncol = 6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "top") + 
  labs(x = "Average neighbourhood size" , y = "Max CT purity\n(across neighbourhoods)" ) +
  theme(axis.text=element_text(size=12)) +
  theme(legend.text=element_text(size=12) , legend.title = element_text(size=12) ) + 
  theme(text=element_text(size=12) , axis.text.x = element_text(size=10) , axis.text.y = element_text(size=12)) + 
  ylim(c(0,1))
p
ggsave(filename = paste0(figures.dir, "avg_hood_size__ct_specificity", ".png"), plot = p, width = 13, height = 7)


# ct enrichment
p = ggplot(tab , aes(x = avg_hood_size_bin , y = avg_celltype_fraction , col = order )) +
  geom_boxplot() + geom_jitter(width = 0.1 , size = 0.2) +
  scale_color_manual(values = cols_order , name = "Order") +
  theme_bw()  + 
  facet_wrap(~celltype_proper, scales = "free",  ncol = 6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "top") + 
  labs(x = "Average neighbourhood size" , y = "Relative CT\nenrichment" ) +
  theme(axis.text=element_text(size=12)) +
  theme(legend.text=element_text(size=12) , legend.title = element_text(size=12) ) + 
  theme(text=element_text(size=12) , axis.text.x = element_text(size=10) , axis.text.y = element_text(size=12)) + 
  ylim(c(0.15,1))
p
ggsave(filename = paste0(figures.dir, "avg_hood_size__average_ct_enrichment", ".png"), plot = p, width = 13, height = 7)



```


# miloDE stat

## Get stat

```{r get-de-stat, message = FALSE}



get_de_stat_single_milo = function(milo_de , celltype){
  current.genes = simulated_genes$ENSEMBL[simulated_genes$celltype == celltype]
  current.milo_de = milo_de[milo_de$ENSEMBL %in% current.genes, ]
  current.milo_de$order_k = paste0(current.milo_de$order, "_" , current.milo_de$k)
  current.milo_de$avg_hood_size = round(current.milo_de$avg_hood_size)
  
  colnames(current.milo_de)[colnames(current.milo_de) == "celltype"] = "celltype_annotated"
  colnames(current.milo_de)[colnames(current.milo_de) == "celltype"] = "celltype_annotated"
  
  order_k = current.milo_de$order_k[1]
  round = current.milo_de$round[1]
  
  current.stat_spec = stat_ct_specificity[stat_ct_specificity$ct_specificity > 0 & stat_ct_specificity$celltype == celltype & stat_ct_specificity$order_k == order_k & stat_ct_specificity$round == round , c("Nhood" , "ct_specificity" , "ct_n_cells", "ct_inclusion")]
  
  current.milo_de = merge(current.milo_de , current.stat_spec , all.x = F , all.y = T)
  current.milo_de$celltype = celltype
  current.milo_de$pval_corrected_across_nhoods[is.na(current.milo_de$pval_corrected_across_nhoods)] = 1
  current.milo_de$pval[is.na(current.milo_de$pval)] = 1
  current.milo_de$gene_sig_bool_pval = current.milo_de$pval < 0.05
  current.milo_de$gene_sig_bool_spatfdr = current.milo_de$pval_corrected_across_nhoods < 0.1
  return(current.milo_de)
}




milo_de_stat = bplapply(1:nrow(anno) , function(i){
  milo_de = readRDS(file = paste0(root.dir ,  "data/processed/order_analysis_full_ds/milo_de_selected/" ,"milo_de_order_", anno$order[i] , "_k_" , anno$k[i] , "_round_" , anno$round[i], ".Rds"))
  
  stat_per_ct = lapply(ct_levels  , function(celltype){
    out = get_de_stat_single_milo(milo_de , celltype)
    return(out)
  })
  stat_per_ct = do.call(rbind , stat_per_ct)
  stat_per_ct$idx = i
  return(stat_per_ct)
} , BPPARAM = mcparam)
milo_de_stat = do.call(rbind , milo_de_stat)


# add additional stat
milo_de_stat = merge(milo_de_stat , stat_ct_specificity_per_coverage , all.x = T)
milo_de_stat = merge(milo_de_stat , simulated_genes[, c("celltype" , "ENSEMBL" , "range" , "frac_2_change" , "logFC_bulk", "logFC_round")] , by = c("celltype" , "ENSEMBL") , all.x = T)
saveRDS(milo_de_stat , paste0(root.dir , "data/processed/order_analysis_full_ds/milo_de_aggregated_stat.Rds"))
milo_de_stat = readRDS(paste0(root.dir , "data/processed/order_analysis_full_ds/milo_de_aggregated_stat.Rds"))


```


# DE detection

## Scater - CT purity, n-cells in CT & DE detection

```{r ct-purity-n-cells-detection-scater, message = FALSE}


stat = as.data.frame(milo_de_stat[abs(milo_de_stat$logFC_bulk) >= 1 & abs(milo_de_stat$logFC_bulk) <= 6 , ] %>% group_by(celltype ,  order_k , order  , round, Nhood) 
                     %>% dplyr::summarise(frac_hoods_sig_pval = mean(gene_sig_bool_pval) , 
                                          frac_hoods_sig_spatfdr = mean(gene_sig_bool_spatfdr),
                                          mean_pval_corrected_across_nhoods = mean(-log10(pval_corrected_across_nhoods)),
                                          ct_n_cells = mean(ct_n_cells) , hood_size = mean(hood_size) , 
                                          ct_specificity = mean(ct_specificity)))
stat = stat[order(stat$frac_hoods_sig_pval) , ]
stat$order = paste0("Order = ", stat$order )


plots = lapply(ct_levels , function(celltype){
  p = ggplot(stat[stat$celltype == celltype , ] , aes(x = ct_specificity , y = ct_n_cells , col = frac_hoods_sig_pval)) +
    geom_point(size = 1) +
    #geom_density_2d(data = df , aes(x = ct_specificity , y = ct_n_cells) , col = "firebrick3") +
    theme_bw() + 
    scale_color_viridis(discrete = F , name = "", breaks = c(0, 0.33 , 0.66 , 1)) + 
    facet_wrap(~order , nrow = 2) + 
    labs(x = "" , y = "") + 
    ggtitle(ct_anno$celltype_proper[ct_anno$celltype == celltype]) + 
    theme(legend.position = "top")  + 
    labs(x = "" , y = "") +
    theme(axis.text=element_text(size=12)) +
    theme(legend.text=element_text(size=12) , legend.title = element_text(size=12) ) + 
    theme(text=element_text(size=12) , axis.text.x = element_text(size=10) , axis.text.y = element_text(size=12))
  return(p)
})
p = ggarrange(plotlist = plots , ncol = 6 , nrow = 4 , common.legend = F)
ggsave(filename = paste0(figures.dir, "scater_de_detection", ".png"), plot = p, width = 19, height = 16)


```

## Boxplot - max detection

```{r ct-purity-n-cells-detection-boxplot, message = FALSE}


stat_per_assignment = as.data.frame(stat %>% group_by(celltype ,  order_k  , round) 
                     %>% dplyr::summarise(max_frac_hoods_sig_pval = max(frac_hoods_sig_pval) ) )

ct_spec.thresh = 0
tab = as.data.frame(stat_ct_specificity %>% group_by(celltype , avg_hood_size , order , k , order_k,  prop , round) %>% dplyr::summarise(n_hoods = sum(ct_specificity > ct_spec.thresh) , max_ct_specificity = max(ct_specificity) , max_ct_inclusion = max(ct_inclusion)))
tab = merge(tab , stat_ct_specificity_per_coverage , all.x = T)
tab = merge(tab , stat_per_assignment, by = c("celltype" , "order_k" , "round"))
tab = merge(tab , ct_anno , all.x = T , all.y = F)
ct_proper_levels = sapply(ct_levels , function(ct) ct_anno$celltype_proper[ct_anno$celltype == ct])
tab$celltype_proper = factor(tab$celltype_proper , ct_proper_levels)

p = ggplot(tab , aes(x = avg_hood_size_bin , y = max_frac_hoods_sig_pval , col = order )) +
  geom_boxplot() + geom_point() +
  scale_color_manual(values = cols_order , name = "Order") +
  theme_bw()  + 
  facet_wrap(~celltype_proper, scales = "free_x" , ncol = 6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1 )) +
  theme(legend.position = "top") + 
  labs(x = "Average neighbourhood size" , y = "Max DE detection power" ) +
  theme(axis.text=element_text(size=12)) +
  theme(legend.text=element_text(size=12) , legend.title = element_text(size=12) ) + 
  theme(text=element_text(size=12) , axis.text.x = element_text(size=10) , axis.text.y = element_text(size=12)) + ylim(c(0.18 , 1.02))
p
ggsave(filename = paste0(figures.dir, "boxplot_max_de_detection", ".png"), plot = p, width = 13, height = 7)






```

# Session Info

```{r sessinf}
sessionInfo()
```
